### 2009-01-08 pkg/docs/SpatialFun/code.Rnw
git-svn-id: svn://svn.r-forge.r-project.org/svnroot/diseasemapping@19 edb9625f-4e0d-4859-8d74-9fd3b1da38cb

library(diseasemapping)
data(casedata)
data(popdata)
head(casedata)
head(popdata@data)
cancerRates = getRates(casedata, popdata, ~age*sex)
summary(cancerRates)
ontario = getSMR(cancerRates, popdata, casedata) 
head(ontario@data)
spplot(ontario, "SMR")

library(glmmBUGS)
data(popDataAdjMat)
library(spdep)
popDataAdjMat = poly2nb(ontario, row.names=as.character(ontario$CSDUID))

forBugs = glmmBUGS(formula = Observed + logExpected ~ 1, effects= "CSDUID", 
                   family="poisson", spatial =popDataAdjMat, data=ontario@data)
names(forBugs$ragged)
startingValues=forBugs$startingValues
source("getInits.R")
library(R2WinBUGS)
ontarioResult = bugs(forBugs$ragged, getInits, parameters.to.save=names(getInits()), 
                     model.file="model.bug", n.chain=3, n.iter=1000, n.burnin=100, 
                     n.thin=10, program="winbugs")

ontarioParams=restoreParams(ontarioResult, forBugs$ragged)
ontarioSummary = summaryChain(ontarioParams)

ontario = mergeBugsData(ontario, ontarioSummary)
head(ontario@data)
spplot(ontario, "RCSDUIDSpatial")

Sweave("code.Rnw")