### 2009-03-26 docs/survival/survival.tex
git-svn-id: svn://svn.r-forge.r-project.org/svnroot/diseasemapping@145 edb9625f-4e0d-4859-8d74-9fd3b1da38cb

\documentclass{article}
\usepackage{amsmath}
\begin{document}
\section{The model}
Person $i$ lives at location $s_i$, and is susceptible to cancer if $Y_i=1$.  If they are susceptible, they have a cancer incidence time $T_i$.

\begin{align*}
Y_i & \sim \text{Bernoulli}(\rho)\\
[T_i | Y_i = 1] &\sim \text{Weibull}(\mu_i, \nu)\\
[T_i | Y_i = 0] & = \infty \\
\log(\mu_i) &= X_i \beta + U(s_i)\\
\text{cov}[U(s_0 + s), U(s_0)] &= \sigma^2 M(s; \theta, \kappa) 
\end{align*}

\section{The Data}
For each individual we observe an enrolment date $D_i$, an event date $E_i$, and an indicator variable $Z_i = 1$ if $E_i$ is a cancer incidence and 0 otherwise.

\[
pr(E_i, Z_i | D_i) = 
\begin{cases}
pr(T_i > E_i | T_i > D_i, Y_i =0) pr(T_i > D_i) pr(Y_i = 0) + pr(Y_i =1) & Z_i = 0 \\
pr(T_i = E_i | T_i > D_i, Y_i =0) pr(T_i > D_i) pr(Y_i = 0) & Z_i = 1
\end{cases}
\]

\end{document}