### 2009-03-26 docs/survival/survival.tex
git-svn-id: svn://svn.r-forge.r-project.org/svnroot/diseasemapping@146 edb9625f-4e0d-4859-8d74-9fd3b1da38cb

\documentclass{article}
\usepackage{amsmath}
\usepackage[margin=1in]{geometry}

\begin{document}
\section{The model}
Person $i, i=1\ldots N$ lives at location $s_i$, and is not susceptible to cancer (cured) if $Y_i=1$.  If they are susceptible, they have a cancer incidence time $T_i$.
\begin{align*}
Y_i & \sim \text{Bernoulli}(\rho)\\
[T_i | Y_i = 0] &\sim \text{Weibull}(\mu_i, \nu)\\
[T_i | Y_i = 1] & = \infty \\
\log(\mu_i) &= X_i \beta + U(s_i)\\
\text{cov}[U(s_0 + s), U(s_0)] &= \sigma^2 M(s; \theta, \kappa) 
\end{align*}
$\mu_i$ and $\nu$ are the scale and shape parameters of the Weibull respectively, and $\rho$ is the cure fraction.  $U(s)$ is a Gaussian random field with a Matern correlation function with scale parameter $\theta$ and roughness $\kappa$.  In practice $U(s)$ will be approximated by a Markov random field $\bar U$ on a grid $g_1 \ldots g_M$ with $U(g_k) \approx \bar U_k$.

\section{The Likelihood}
For each individual we observe an enrolment date $L_i$, an event (or death) date $R_i$, and an indicator variable $Z_i = 1$ if $R_i$ is a cancer incidence and 0 if it's a censoring time. The density of a Weibull with scale parameter $\mu$ and shape parameter $\nu$ is 
\[
f(x;\mu,\nu) = \frac{\nu}{\mu} \left(\frac{x}{\mu}\right)^{\nu-1} \exp\left[-\left(\frac{x}{\mu}\right)^\nu\right],
\]
and
\[
\int_0^x f(u;\mu,\nu)du = 1-\exp\left[-\left(\frac{x}{\mu}\right)^\nu\right].
\]
using Wikipedia's parametrization.

Censored observations:
\begin{align*}
pr(R_i, Z_i=0 | L_i) =& pr(T_i > R_i | T_i > L_i, Y_i =0) pr(Y_i =0)  + pr(Y_i =1)\\
=&pr(T_i > R_i) pr(Y_i = 0)/pr(T_i > L_i) + pr(Y_i =1)\\
=& \rho + (1-\rho) \left. \int_{R_i}^\infty f(x;\mu_i, \nu) dx \right/\int_{L_i}^\infty f(x;\mu_i, \nu) dx  \\
= &  \rho + (1-\rho) \left. \exp\left[-\left(\frac{R_i}{\mu_i}\right)^\nu\right] \right/\exp\left[-\left(\frac{L_i}{\mu_i}\right)^\nu\right]\\
= &  \rho + (1-\rho) \exp\left[\left(\frac{L_i}{\mu_i}\right)^\nu -\left(\frac{R_i}{\mu_i}\right)^\nu\right] \\
\end{align*}

Observed events:
\begin{align*}
pr(R_i, Z_i=1 | L_i) =& pr(T_i = R_i | T_i > L_i, Y_i =0)  pr(Y_i = 0) 
\\
=& 
 pr(T_i = L_i) pr(Y_i = 0) /pr(T_i > L_i) \\
=&
 (1-\rho) f(L_i ; \mu_i, \nu) \left/ \int_{L_i}^\infty f(x;\mu_i, \nu) dx  \right.\\
=&
 (1-\rho) \frac{\nu}{\mu_i} \left(\frac{R_i}{\mu_i}\right)^{\nu-1} \exp\left[\left(\frac{L_i}{\mu_i}\right)^\nu-\left(\frac{R_i}{\mu_i}\right)^\nu\right]
\end{align*}

\end{document}