### 2009-04-02 docs/inla/inla.rnw
git-svn-id: svn://svn.r-forge.r-project.org/svnroot/diseasemapping@168 edb9625f-4e0d-4859-8d74-9fd3b1da38cb

\documentclass[xcolor=dvipsnames]{beamer}

\usetheme{Singapore}
\usecolortheme[named=RawSienna]{structure}

\usefonttheme{serif}
\setbeamertemplate{navigation symbols}{}
\setbeamertemplate{footline}[frame number]

\usepackage[nogin]{c:/brownpe/share/Sweave}
\SweaveOpts{echo=FALSE,fig=true,prefix.string=Figures/G,height=2,width=3}

\graphicspath{{Figures/}}
\setkeys{Gin}{width=\textwidth}

\usepackage{amsmath}
\usepackage{c:/brownpe/share/Sweave}

\newcommand{\comment}[1]{}

\title{Generalized Linear Mixed Models}
\author{Patrick Brown}
\institute{Cancer Care Ontario, and \\ School of Public Health, University of Toronto}

\begin{document}
<<junk,fig=false>>=
options(SweaveHooks=list(fig=function() par(mar=c(2.5,2.5,0.1,0.1), mgp=c(1.5, 0.5, 0), cex=0.6)))
@

\frame{\maketitle}

\begin{frame}
\frametitle{Outline}
\begin{itemize}
	\item The model and the problem
	\item First generation of approximate methods and PQL
	\item Bayesian inference and Markov Chain Monte Carlo
	\item Integrated Nested Laplace Approximations --- the holy grail?
\end{itemize}
\end{frame}

\begin{frame}
\frametitle{The model}
$Y_{ij}$ is the disease status of person $i$ in family $j$.
\begin{align*}
Y_{ij} \sim  & \text{Bernoulli}(p_{ij})\\\
logit(p_{ij}) = &x_{ij}\beta + V_{i}\\
V_{i} \sim &\text{N}(0, \sigma^2)
\end{align*}
$V_i$ is a family-level random effect.

\begin{itemize}
	\item What are the parameter estimates $\hat\sigma^2$ and $\hat\beta$?
	\item and their standard errors?
	\item And what are the predictors of family-level random effects $E(V_i | data)$, and their conditional distributions $[V_i | data]$.
\end{itemize}

\end{frame}

\begin{frame}[fragile]
\frametitle{Example: }
\begin{columns}
\column{0.5\textwidth}
\begin{itemize}
	\item Repeated measurements on the presence of a bacteria in children
	\item Placebo and two different treatments
\end{itemize}

<<thedata1,echo=true,fig=false>>=
library(MASS)
data(bacteria)
head(bacteria)
@

\column{0.5\textwidth}
\begin{itemize}
	\item $Y_{ij}$ is the bacteria presence in child $i$ at time $j$
	\item $X_{ij}$ is a vector of indicator variables for treatment
	\item $V_i$ is each child's relative risk
\end{itemize}
\end{columns}
\end{frame}

\begin{frame}
\frametitle{Example: lung cancer in Ontario}
\begin{columns}
\column{0.65\textwidth}
\begin{itemize}
	\item $Y_i$ and $E_i$ are the observed and expected lung cancer cases in region $i$
	\item $X_i$ is income, a covariate
	\item $U_i$ and $V_i$ are spatial and independent random effects
\end{itemize}

\begin{align*}
Y_i \sim & \text{Posson}(\lambda_i E_i)\\
\log(\lambda_i) = & \mu + X_i \beta + U_i + V_i\\
U_i \sim &MRF(\sigma^2_u)\\
V_i \sim & N(0, \sigma^2_v)
\end{align*}

\column{0.35\textwidth}

Expected 
\includegraphics{ontarioE.jpg}
\comment{
library(diseasemapping)
library(sp)
data(popdata)
data(casedata)
model = getRates(casedata, popdata, ~age*sex)
ontario = getSMR(popdata,model,  casedata)
jpeg('Figures/ontarioE.jpg', height=600, width=800)
print(spplot(ontario, 'logExpected', lty=0))
dev.off()
}

Observed
\includegraphics{ontarioO.jpg}

\comment{
	jpeg('Figures/ontarioO.jpg', height=600, width=800)
 print(spplot(ontario, 'observed',lty=0))
dev.off()
}

\end{columns}

\end{frame}

\begin{frame}
\frametitle{Likelihood function?}
\begin{itemize}
	\item The $V_i$ are not parameters, and not observed, so they must be integrated out of the likelihood:
\end{itemize}
\[
L(\beta, \sigma) = \prod_i \int pr(Y_{ij}|V_i) pr(V_i) dV_i 
\]
\begin{itemize}
	\item The integral is intractable except in a few special cases
	\item such as Normally-distributed $Y_{ij}$ 
	\item or Gamma $Y_{ij}$ and $V_{i}$ with no covariates
	\item Empirical Bayes methods are based on conjugate priors
\end{itemize}
\end{frame}

\begin{frame}
\frametitle{First generation of approximations}
\begin{block}{Can we use least squares?}
\begin{itemize}
	\item With no random effects, model fitting is done with Iterated Reweighted Least Squares, or Fisher Scoring
	\item For Poisson data, pretend $Y_i \sim N(\lambda_i, \lambda_i)$ and use least squares.  Though you need to know the variance to do this.
	\item Choose a starting value, then iterate between calculating the variances and estimating the parameters.
\end{itemize}
\end{block}

\begin{block}{Penalized Quazi-likelihood}
Start with initial values for $V_i$ and $\sigma$, then use normal-based likelihood and Fisher scores to:
\begin{itemize}
	\item estimate $\hat\beta$
	\item update predictions of  $V_i$, 
	\item Re-estimate $\sigma$
	\item repeat
\end{itemize}

\end{block}

\end{frame}

\begin{frame}
\frametitle{Example: Bacteria}

\end{frame}

\begin{frame}
\frametitle{Markov Chain Monte Carlo}

\end{frame}

\begin{frame}
\frametitle{Example:}

\end{frame}

\begin{frame}
\frametitle{Example: Lung cancer in Ontario}

\end{frame}

\begin{frame}
\frametitle{Integrated Nested Laplace Approximations}

\end{frame}

\begin{frame}
\frametitle{How it works}

\end{frame}

\begin{frame}
\frametitle{Example: }

\end{frame}

\begin{frame}
\frametitle{Example: Lung cancer in Ontario}

\end{frame}

\end{document}