### 2009-04-02 docs/inla/inla.rnw
git-svn-id: svn://svn.r-forge.r-project.org/svnroot/diseasemapping@174 edb9625f-4e0d-4859-8d74-9fd3b1da38cb

\documentclass[xcolor=dvipsnames]{beamer}

\usetheme{Singapore}
\usecolortheme[named=RawSienna]{structure}

\usefonttheme{serif}
\setbeamertemplate{navigation symbols}{}
\setbeamertemplate{footline}[frame number]

\usepackage[nogin]{c:/brownpe/share/Sweave}
\SweaveOpts{echo=FALSE,fig=true,prefix.string=Figures/G,height=2,width=3}

\graphicspath{{Figures/}}
\setkeys{Gin}{width=\textwidth}

\usepackage{amsmath}
\usepackage{c:/brownpe/share/Sweave}

\newcommand{\comment}[1]{}

\title{Generalized Linear Mixed Models}
\author{Patrick Brown}
\institute{Cancer Care Ontario, and \\ School of Public Health, University of Toronto}

\begin{document}
<<junk,fig=false>>=
options(SweaveHooks=list(fig=function() par(mar=c(2.5,2.5,0.1,0.1), mgp=c(1.5, 0.5, 0), cex=0.6)))
@

\frame{\maketitle}

\begin{frame}
\frametitle{Outline}
\begin{itemize}
	\item The model and the problem
	\item First generation of approximate methods and PQL
	\item Bayesian inference and Markov Chain Monte Carlo
	\item Integrated Nested Laplace Approximations --- the holy grail?
\end{itemize}
\end{frame}

\begin{frame}
\frametitle{The model}
$Y_{ij}$ is the disease status of person $i$ in family $j$.
\begin{align*}
Y_{ij} \sim  & \text{Bernoulli}(p_{ij})\\\
logit(p_{ij}) = &x_{ij}\beta + V_{i}\\
V_{i} \sim &\text{N}(0, \sigma^2)
\end{align*}
$V_i$ is a family-level random effect.

\begin{itemize}
	\item What are the parameter estimates $\hat\sigma^2$ and $\hat\beta$?
	\item and their standard errors?
	\item And what are the predictors of family-level random effects $E(V_i | data)$, and their conditional distributions $[V_i | data]$.
\end{itemize}

\end{frame}

\begin{frame}[fragile]
\frametitle{Example: }
\begin{columns}
\column{0.5\textwidth}
\begin{itemize}
	\item Repeated measurements on the presence of a bacteria in children
	\item Placebo and two different treatments
\end{itemize}

<<thedata1,fig=false>>=
library(MASS)
data(bacteria)
bacterianew <- bacteria[,-(2:3)]
bacterianew$y = as.integer(bacteria$y=="y")
levels(bacterianew$trt) <- c("placebo", "drug", "drugplus")
bacteria=bacterianew

head(bacteria)
#$
@

\column{0.5\textwidth}
\begin{itemize}
\item the ID column denotes repeated measurements on each child.
	\item $Y_{ij}$ is the bacteria presence in child $i$ at time $j$
	\item $X_{ij}$ is a vector of indicator variables for treatment
	\item $V_i$ is each child's relative risk
\end{itemize}
\end{columns}
\end{frame}

\begin{frame}
\frametitle{Example: lung cancer in Ontario}
\begin{columns}
\column{0.65\textwidth}
\begin{itemize}
	\item $Y_i$ and $E_i$ are the observed and expected lung cancer cases in region $i$
	\item $X_i$ is income, a covariate
	\item $U_i$ and $V_i$ are spatial and independent random effects
\end{itemize}

\begin{align*}
Y_i \sim & \text{Posson}(\lambda_i E_i)\\
\log(\lambda_i) = & \mu + X_i \beta + U_i + V_i\\
U_i \sim &MRF(\sigma^2_u)\\
V_i \sim & N(0, \sigma^2_v)
\end{align*}

\column{0.35\textwidth}

Expected 
\includegraphics{ontarioE.jpg}
\comment{
library(diseasemapping)
library(sp)
data(popdata)
data(casedata)
model = getRates(casedata, popdata, ~age*sex)
ontario = getSMR(popdata,model,  casedata)
jpeg('Figures/ontarioE.jpg', height=600, width=800)
print(spplot(ontario, 'logExpected', lty=0))
dev.off()
}

Observed
\includegraphics{ontarioO.jpg}

\comment{
	jpeg('Figures/ontarioO.jpg', height=600, width=800)
 print(spplot(ontario, 'observed',lty=0))
dev.off()
}

\end{columns}

\end{frame}

\begin{frame}
\frametitle{Likelihood function?}
\begin{itemize}
	\item The $V_i$ are not parameters, and not observed, so they must be integrated out of the likelihood:
\end{itemize}
\[
L(\beta, \sigma) = \prod_i \int pr(Y_{ij}|V_i) pr(V_i) dV_i 
\]
\begin{itemize}
	\item The integral is intractable except in a few special cases
	\item such as Normally-distributed $Y_{ij}$ 
	\item or Gamma $Y_{ij}$ and $V_{i}$ with no covariates
	\item Empirical Bayes methods are based on conjugate priors
\end{itemize}
\end{frame}

\begin{frame}
\frametitle{First generation of approximations}

\begin{itemize}
	\item With no random effects, model fitting is done with Iterated Reweighted Least Squares, or Fisher Scoring
	\item For Poisson data, pretend $Y_i \sim N(\lambda_i, \lambda_i)$ and use least squares.  Though you need to know the variance to do this.
	\item Choose a starting value, then iterate between calculating the variances and estimating the parameters.
\end{itemize}

\begin{block}{Penalized Quazi-likelihood}
Start with initial values for $V_i$ and $\sigma$, then use normal-based likelihood and Fisher scores to:
\begin{itemize}
	\item estimate $\hat\beta$
	\item update predictions of  $V_i$, 
	\item Re-estimate $\sigma$
	\item repeat
\end{itemize}

\end{block}

\end{frame}

\begin{frame}[fragile]
\frametitle{Example: Bacteria}
{\scriptsize
<<bacteriaPQL,fig=false,echo=true>>=
library(MASS)
pql = glmmPQL(y ~ trt, random = ~1|ID, family='binomial',data=bacteria)
summary(pql)
@
}

\end{frame}

\begin{frame}
\frametitle{Advantages/Disadvantages}

\begin{block}{Advantages}
\begin{itemize}
	\item Quick and easy to use
	\item It's better than ignoring dependence completely and using a glm.
\end{itemize}
\end{block}

\begin{block}{Disadvantages}
\begin{itemize}
	\item It doesn't always produce good results
	\item Especially if your data are very non-normal,  with small numbers in each group.
	\item Theoretical justification isn't strong
\end{itemize}
\end{block}

\end{frame}

\begin{frame}
\frametitle{Markov Chain Monte Carlo}
\begin{itemize}
	\item Put prior distributions on parameters $[\sigma]$, $[\beta]$.
	\item Simulate from the posterior distributions
	\[
	[\sigma|Y],\ [\beta|Y],\ [V|Y]
	\]
	\item Use $[V|Y] = [Y|V][V]/[Y]$, but $[Y]$ is impossible to evaluate.
	\item Gibbs sampler: alternate between sampling from $[V|Y,\beta,\sigma]$ and $[\beta,\sigma|V,Y]$.
	\item Random walk Metropolis: update $V_i^{(p)}$ to  $V_i^{(p+1)}~N(V_i^{(p)}, \nu^2)$, accept the move with probability
	\[
	\min[1, pr(Y|V_i^{(p+1)}) /  pr(Y|V_i^{(p)})) ]
	\]
	\item This is guaranteed to converge to a sample from $[V_i|Y]$\ldots
	\item \ldots eventually
\end{itemize}
\end{frame}

\begin{frame}[fragile]
\frametitle{Example:bacteria}
{\scriptsize
<<bacmcmc,fig=false,echo=true>>=
library(glmmBUGS)
bac = glmmBUGS(y ~ trt, effects='ID', family='bernoulli',data=bacteria, modelFile='modelx.bug')
startingValues = bac$startingValues
source("getInits.R")
library(BRugs)
library(R2WinBUGS)
bacResult = bugs(bac$ragged, getInits, 
	parameters.to.save = names(getInits()),
  n.chain=3, n.iter=100, n.burnin=10, n.thin=2,
  program="openbugs", model.file='modelBac.bug') 
bacParams = restoreParams(bacResult, bac$ragged)

@
}
\end{frame}

\begin{frame}[fragile]
\frametitle{Posterior samples}
<<bacSample,echo=true>>=
hist(bacParams$SDID)
@
\end{frame}

\begin{frame}[fragile]
\frametitle{Parameter estimates}
<<bacpar,fig=false,echo=true>>=
bacSummary = summaryChain(bacParams)
bacSummary$scalars[,c(1:3,5,8)]
bacSummary$RID[1:4,c(1:3,5,8)]

@
\end{frame}

\begin{frame}[fragile]
\frametitle{Did the chiain converge?}
<<bacchain,echo=true>>=
checkChain(bacParams)
@

\end{frame}

\begin{frame}[fragile]
\frametitle{Example: Lung cancer in Ontario}
\begin{verbatim}
popDataAdjMat = poly2nb(ontario,row.names=as.character(ontario[["CSDUID"]]))

forBugs = glmmBUGS(formula=observed + logExpected ~ 1,
  effects="CSDUID",   family="poisson", 
  spatial=popDataAdjMat,
  data=ontario@data)
startingValues = forBugs$startingValues
source("getInits.R")

ontarioResult = bugs(forBugs$ragged, getInits, 
  parameters.to.save = names(getInits()),
    model.file="model.bug", n.chain=3, n.iter=50, 
    n.burnin=10, n.thin=2,
      program="winbugs", debug=T)

ontarioParams = restoreParams(ontarioResult, forBugs$ragged)
\end{verbatim}
\end{frame}

\comment(

\comment{
	jpeg('Figures/ontarioCancer.jpg', height=600, width=800)
}

\begin{frame}[fragile]
\frametitle{Predicted cancer rates $E(U_i + V_i|Y)$}
\begin{verbatim}
ontarioSummary = summaryChain(ontarioParams)
ontario = mergeBugsData(popdata, ontarioSummary)
spplot(ontario, "RCSDUID.mean")
\end{verbatim}
\includegraphics[width=0.8\textwidth]{ontarioCancer.jpg}
\end{frame}

\begin{frame}
\frametitle{Integrated Nested Laplace Approximations}

\end{frame}

\begin{frame}
\frametitle{How it works}

\end{frame}

\begin{frame}
\frametitle{Example: }
<<bacteriaINlA,fig=false>>=
#library(MASS)
#data(bacteria)
#bacterianew <- bacteria[,-(2:3)]
#bacterianew$y = as.integer(bacteria$y=="y")
#levels(bacterianew$trt) <- c("placebo", "drug", "drugplus")
#bacteria=bacterianew
# do this pql = glmmPQL(y ~ trt, random = ~1|ID, family='binomial',data=bacteria)

@

\end{frame}

\begin{frame}
\frametitle{Example: Lung cancer in Ontario}
<<cancerInla,fig=false>>=
if(F) {
library(diseasemapping)
library(sp)
data(popdata)
data(casedata)
model = getRates(casedata, popdata, ~age*sex)
ontario = getSMR(popdata,model,  casedata)
data(popDataAdjMat)

# fit bym model with inla, like
#forBugs = glmmBUGS(formula=Observed + logExpected ~ 1,
#  effects="CSDUID",   family="poisson", spatial=popDataAdjMat,
#  data=ontario)
}
@
\end{frame}

\end{document}