### 2009-04-06 docs/dionne/slides.rnw
git-svn-id: svn://svn.r-forge.r-project.org/svnroot/diseasemapping@179 edb9625f-4e0d-4859-8d74-9fd3b1da38cb

\documentclass[xcolor=dvipsnames]{beamer}

\usetheme{Singapore}
\usecolortheme[named=RawSienna]{structure}

\usefonttheme{serif}
\setbeamertemplate{navigation symbols}{}
\setbeamertemplate{footline}[frame number] 
\usepackage[nogin]{c:/brownpe/share/Sweave}
\title{Modelling small-area disease incidence data in space and time }
\author{Patrick Brown}
\institute{Cancer Care Ontario, and \\ School of Public Health, University of Toronto}

\usepackage{amsmath}
\graphicspath{{Figures/}}
\setkeys{Gin}{width=\textwidth}
\SweaveOpts{echo=FALSE,fig=true,prefix.string=Figures/G,height=2,width=3}

\begin{document}

\frame{\maketitle}

\begin{frame}
\frametitle{Large v Small Area Maps}

\begin{columns}
\column{0.5\textwidth}

\begin{block}{Coarse scale}

\vspace{10pt}

\includegraphics[width=0.9\textwidth]{lung.png}

\vspace{10pt}

Observed rate for Ontario, by Census sub-division

\end{block}

\column{0.5\textwidth}

\begin{block}{Fine scale}

\includegraphics[width=\textwidth]{torCTsmr.pdf}

SMR's for Toronto, by census tract

\end{block}

\end{columns}

\end{frame}

\begin{frame}
\frametitle{Issues with small area maps}

\begin{block}{Small counts}
	mostly 0 or 1
\begin{itemize}
\item The usual estimates of incidence rates perform poorly
	\item Methods based on the normal-distribution aren't appropriate
\end{itemize}

\end{block}
\begin{block}{Spatial dependence}
With regions close together being more similar than regions further apart
\begin{itemize}
	\item The usual regression models underestimate standard errors
  \item but this helps us deal with small counts
\end{itemize}

\end{block}

\end{frame}

\begin{frame}
\frametitle{This talk}
Following this talk, those here will be able to:
\begin{itemize}
\item Understand a statistical model suitable for small-area disease incidence data with low counts;
\item Use the results of this model to answer questions in social and environmental epidemiology;
\item  Be aware of how to extend these models to problems in space and time.
\end{itemize}
\end{frame}

\begin{frame}
\frametitle{ North Carolina}
describe data
map raw rates
map covariates, \% black, unemp
\end{frame}

\begin{frame}
\frametitle{The model (without equations)}

\begin{itemize}
	\item The true risk surface $\lambda(s)$ at location $s$ is unknown.  
	\item It depends on some measured risk factors, such as income and unemployment rate $X(s)$.
	\item And some residual spatial variation $W(s)$ which might be smoothly varying in space.

\begin{itemize}
	\item This might be caused by unmeasured risk factor such as air pollution
	\item Or be the result of direct transmission creating areas of high risk
\end{itemize}
\item The number of cases $Y_i$ in region $i$ is random, but we'd expect it to be related to the population in the region and the risk $\lambda(s)$ there.
\item In regions of small population we shouldn't be surprised to see zero cases, even if risk isn't zero.
\end{itemize}

\end{frame}

\begin{frame}
\frametitle{Dealing with small counts}

\begin{columns}
\column{0.5\textwidth}
\begin{block}{The Model}
\[
Y_i \sim \text{Poisson}(P_i\lambda_i)
\]
\begin{itemize}
	\item $Y_i$ is the number of cases in region $i$
	\item $P_i$ is the population
	\item $\lambda_i$ is the risk
\end{itemize}
\end{block}
\column{0.5\textwidth}
\begin{block}{The Poisson Distriution}
<<poisson,fig=true,echo=false>>=
options(SweaveHooks=list(fig=function() par(mar=c(2.5,2.5,0.1,0.1), mgp=c(1.5, 0.5, 0), cex=0.6)))
xseq = 0:14
toplot = dpois(xseq, 0.5)
names(toplot) = as.character(xseq)
toplot2 = dpois(xseq, 1.5)
barplot(toplot, width=0.4, xlim = range(xseq), space=4, xlab='count', ylab='pr(count)',
	col='red',border=NA)
barplot(toplot, width=0.21, add=T, space=8.8, col='white', xlim = range(xseq),border=NA, axes=F, axisnames=F)

barplot(toplot2, width=0.2, add=T, space=9, col='blue', xlim = range(xseq),border=NA, axes=F)
@

mean = \textcolor{red}{0.5},  \textcolor{blue}{1.5} 
\end{block}

\end{columns}

\begin{block}{Interpretation}
\begin{itemize}
	\item We don't observe the true rate $\lambda_i$
	\item but rather a noisy version of it in the form of case counts
	\item The Poisson distribution can cope with small counts and lots of zeros
\end{itemize}
\end{block}

\end{frame}

\begin{frame}
\frametitle{Modelling the rates}
\begin{align*}
Y_i \sim & \text{Poisson}(P_i\lambda_i)\\
\log(\lambda_i) = & \mu + X_i \beta + U_i + V_i\\
V_i \sim &\text{N}(0, \tau^2)\\
U_i|U_{-i} \sim & \text{N}(\text{mean}\{U_j; j \sim i \}, \sigma^2/||\{j ; j \sim i\}||
\end{align*}
\begin{itemize}
  \item Random effects model: Risks are usually close to the average --- shrinkage
	\item Risk can depends on covariates $X_i$ (i.e.\ income)
	\item $U_i$ is a spatially smooth random effect, $V_i$ is a rough non-spatial effect
\item The spatial nature of $U_i$ results in regions neighbouring $i$ to influence risk estimation at $i$, somewhat mitigating the low counts problem.
\end{itemize}

\end{frame}

\begin{frame}
\frametitle{Model fitting}
\begin{itemize}
	\item The data are not Normally-distributed, and there is an unobserved random effect.
	\item This makes the usual least squares or maximum likelihood methods problematic
	\item Bayesian inference using Markov Chain Monte Carlo simulation
	\item There are several implementations in the statistical package R.
	\item and the ArcGIS add-on the Rapid Inquiry Facility
\end{itemize}
\end{frame}

\begin{frame}
\frametitle{Example: North Carolina}

\begin{block}{Where:
Predicted risk E$(\lambda_i | data)$}

\includegraphics[width=0.8\textwidth]{rate.png}
\end{block}

\begin{block}{Why:
Predicted risk minus known risk factors}
$E(\lambda_i - X_i \beta | data)= E(U_i + V_i)$

\includegraphics[width=0.8\textwidth]{relativeRate.png}

\end{block}

\end{frame}

\begin{frame}
\vspace*{-15pt}
Parameter estimates\\
\vspace*{-5pt}
{\small
<<parest,results=tex,fig=false>>=
if(F) {
load("data\\GONQResult.RData")
library(glmmBUGS)
  gonpar = restoreParams(GONQResult, GONQragged$ragged)
  save(gonpar, file="gonpar.RData")
}
load("gonpar.RData")
library(glmmBUGS)
gonSummary = summaryChain(gonpar)
par = rbind(gonSummary[[1]], gonSummary[[2]])

toprint =par[c(1,3,4,16:32),c(1,2,5,9)]
rownames(toprint) = gsub("_", "-" , rownames(toprint))
colnames(toprint) = gsub("%", "pct",colnames(toprint))
latex(toprint, file='', dec=2, rowlabel='Parameter')
@
}
\end{frame}

\begin{frame}
\frametitle{Recap of spatial models:}
\begin{itemize}
	\item True risk is unobserved, use observed counts to predict risk
	\item Risk is random and may vary smoothly in space
\item Allow for known risk factors, as covariates
\item Map residual spatial variation or predicted risk
\item and look at parameter estimates
\end{itemize}
\end{frame}

\begin{frame}
\frametitle{Spatio-temporal models}
\begin{itemize}
	\item Risk changes in time as well as space
	\item $Y_{ij}$ is the incidence count in region $i$ at time $j$
	\begin{align*}
Y_{ij} \sim & \text{Poisson}(P_{ij}\lambda_{ij})\\
\log(\lambda_{ij}) = & \mu + X_{ij} \beta + U_i + V_i + W_{ij}\\
V_i \sim &\text{N}(0, \tau^2)\\
U_i|U_{-i} \sim & \text{N}(\text{mean}\{U_j; j \sim i \}, \sigma^2/||\{j ; j \sim i\}||\\
W_{ij} \sim &\text{N}(0, \nu^2)
	\end{align*}
\item $W_{ij}$ is the variation in time from the long-term average for each region.
\end{itemize}

\end{frame}

\begin{frame}
\frametitle{Parameter Estimates}
<<parestst,results=tex,fig=false>>=

toprint =par[2:12,c(1,2,5,9)]
rownames(toprint) = gsub("_", "-" , rownames(toprint))
colnames(toprint) = gsub("%", "pct",colnames(toprint))
latex(toprint, file='', dec=2, rowlabel='Parameter')
@

\end{frame}

\begin{frame}
\frametitle{E($W_{ij}$), time 1}
\includegraphics[width=1.2\textwidth]{R05Q2.png}
\end{frame}

\begin{frame}
\frametitle{E($W_{ij}$), time 2}
\includegraphics[width=1.2\textwidth]{R05Q3.png}
\end{frame}

\begin{frame}
\frametitle{E($W_{ij}$), time 3}
\includegraphics[width=1.2\textwidth]{R05Q4.png}
\end{frame}

\begin{frame}
\frametitle{E($W_{ij}$), time 4}
\includegraphics[width=1.2\textwidth]{R06Q1.png}
\end{frame}

\begin{frame}
\frametitle{E($W_{ij}$), time 5}
\includegraphics[width=1.2\textwidth]{R06Q2.png}
\end{frame}
\begin{frame}
\frametitle{E($W_{ij}$), time 6}
\includegraphics[width=1.2\textwidth]{R06Q3.png}
\end{frame}
\begin{frame}
\frametitle{E($W_{ij}$), time 7}
\includegraphics[width=1.2\textwidth]{R06Q4.png}
\end{frame}
\begin{frame}
\frametitle{E($W_{ij}$), time 8}
\includegraphics[width=1.2\textwidth]{R07Q1.png}
\end{frame}
\begin{frame}
\frametitle{E($W_{ij}$), time 9}
\includegraphics[width=1.2\textwidth]{R07Q2.png}
\end{frame}
\begin{frame}
\frametitle{E($W_{ij}$), time 10}
\includegraphics[width=1.2\textwidth]{R07Q3.png}
\end{frame}
\begin{frame}
\frametitle{E($W_{ij}$), time 11}
\includegraphics[width=1.2\textwidth]{R07Q4.png}
\end{frame}
\begin{frame}
\frametitle{E($W_{ij}$), time 12}
\includegraphics[width=1.2\textwidth]{R08Q1.png}
\end{frame}

\begin{frame}
\frametitle{Conclusion}

\begin{itemize}
	\item Small area disease mapping in space and time is possible even when counts are low 
	\item Include covariates in the model to estimate the effect of potential risk factors
	\item There is software for implementing these models which is not terribly difficult to use.
\end{itemize}
\end{frame}

\end{document}