### 2009-08-21 docs/hamilton/graph.tex
git-svn-id: svn://svn.r-forge.r-project.org/svnroot/diseasemapping@362 edb9625f-4e0d-4859-8d74-9fd3b1da38cb

% C:\brownpe\research\diseasemapping\docs\hamilton\Figures\graph.jdr
% Created by Jpgfdraw version 0.5.4b
% Aug 18, 2009 9:28:53 AM
\begin{pgfpicture}{0bp}{0bp}{347.0bp}{247.5bp}
\begin{pgfscope}
\pgfsetlinewidth{3.0bp}
\pgfsetrectcap 
\pgfsetmiterjoin \pgfsetmiterlimit{10.0}
\pgfpathmoveto{\pgfpoint{93.75bp}{213.5bp}}
\pgfpathcurveto{\pgfpoint{93.75bp}{196.103031bp}}{\pgfpoint{76.624893bp}{182.0bp}}{\pgfpoint{55.5bp}{182.0bp}}
\pgfpathcurveto{\pgfpoint{34.375109bp}{182.0bp}}{\pgfpoint{17.25bp}{196.103031bp}}{\pgfpoint{17.25bp}{213.5bp}}
\pgfpathcurveto{\pgfpoint{17.25bp}{230.89697bp}}{\pgfpoint{34.375109bp}{245.0bp}}{\pgfpoint{55.5bp}{245.0bp}}
\pgfpathcurveto{\pgfpoint{76.624893bp}{245.0bp}}{\pgfpoint{93.75bp}{230.89697bp}}{\pgfpoint{93.75bp}{213.5bp}}
\pgfclosepath
\color[rgb]{0.0,0.0,0.0}
\pgfusepath{stroke}
\end{pgfscope}
\begin{pgfscope}
\pgfsetlinewidth{3.0bp}
\pgfsetrectcap 
\pgfsetmiterjoin \pgfsetmiterlimit{10.0}
\pgfpathmoveto{\pgfpoint{318.5bp}{214.5bp}}
\pgfpathcurveto{\pgfpoint{318.5bp}{197.103031bp}}{\pgfpoint{301.374893bp}{183.0bp}}{\pgfpoint{280.25bp}{183.0bp}}
\pgfpathcurveto{\pgfpoint{259.125109bp}{183.0bp}}{\pgfpoint{242.0bp}{197.103031bp}}{\pgfpoint{242.0bp}{214.5bp}}
\pgfpathcurveto{\pgfpoint{242.0bp}{231.89697bp}}{\pgfpoint{259.125109bp}{246.0bp}}{\pgfpoint{280.25bp}{246.0bp}}
\pgfpathcurveto{\pgfpoint{301.374893bp}{246.0bp}}{\pgfpoint{318.5bp}{231.89697bp}}{\pgfpoint{318.5bp}{214.5bp}}
\pgfclosepath
\color[rgb]{0.0,0.0,0.0}
\pgfusepath{stroke}
\end{pgfscope}
\begin{pgfscope}
\pgfsetlinewidth{3.0bp}
\pgfsetrectcap 
\pgfsetmiterjoin \pgfsetmiterlimit{10.0}
\pgfpathmoveto{\pgfpoint{181.25bp}{105.75bp}}
\pgfpathcurveto{\pgfpoint{181.25bp}{88.353031bp}}{\pgfpoint{164.124893bp}{74.25bp}}{\pgfpoint{143.0bp}{74.25bp}}
\pgfpathcurveto{\pgfpoint{121.875109bp}{74.25bp}}{\pgfpoint{104.75bp}{88.353031bp}}{\pgfpoint{104.75bp}{105.75bp}}
\pgfpathcurveto{\pgfpoint{104.75bp}{123.14697bp}}{\pgfpoint{121.875109bp}{137.25bp}}{\pgfpoint{143.0bp}{137.25bp}}
\pgfpathcurveto{\pgfpoint{164.124893bp}{137.25bp}}{\pgfpoint{181.25bp}{123.14697bp}}{\pgfpoint{181.25bp}{105.75bp}}
\pgfclosepath
\color[rgb]{0.0,0.0,0.0}
\pgfusepath{stroke}
\end{pgfscope}
\begin{pgfscope}
\pgfsetlinewidth{3.0bp}
\pgfsetrectcap 
\pgfsetmiterjoin \pgfsetmiterlimit{10.0}
\pgfpathmoveto{\pgfpoint{266.75bp}{107.25bp}}
\pgfpathcurveto{\pgfpoint{266.75bp}{89.853031bp}}{\pgfpoint{249.624893bp}{75.75bp}}{\pgfpoint{228.5bp}{75.75bp}}
\pgfpathcurveto{\pgfpoint{207.375109bp}{75.75bp}}{\pgfpoint{190.25bp}{89.853031bp}}{\pgfpoint{190.25bp}{107.25bp}}
\pgfpathcurveto{\pgfpoint{190.25bp}{124.64697bp}}{\pgfpoint{207.375109bp}{138.75bp}}{\pgfpoint{228.5bp}{138.75bp}}
\pgfpathcurveto{\pgfpoint{249.624893bp}{138.75bp}}{\pgfpoint{266.75bp}{124.64697bp}}{\pgfpoint{266.75bp}{107.25bp}}
\pgfclosepath
\color[rgb]{0.0,0.0,0.0}
\pgfusepath{stroke}
\end{pgfscope}
\begin{pgfscope}
\pgfsetlinewidth{3.0bp}
\pgfsetrectcap 
\pgfsetmiterjoin \pgfsetmiterlimit{10.0}
\pgfpathmoveto{\pgfpoint{166.5bp}{29.75bp}}
\pgfpathcurveto{\pgfpoint{166.5bp}{14.838318bp}}{\pgfpoint{129.563492bp}{2.75bp}}{\pgfpoint{84.0bp}{2.75bp}}
\pgfpathcurveto{\pgfpoint{38.436508bp}{2.75bp}}{\pgfpoint{1.5bp}{14.838318bp}}{\pgfpoint{1.5bp}{29.75bp}}
\pgfpathcurveto{\pgfpoint{1.5bp}{44.661682bp}}{\pgfpoint{38.436508bp}{56.75bp}}{\pgfpoint{84.0bp}{56.75bp}}
\pgfpathcurveto{\pgfpoint{129.563492bp}{56.75bp}}{\pgfpoint{166.5bp}{44.661682bp}}{\pgfpoint{166.5bp}{29.75bp}}
\pgfclosepath
\color[rgb]{0.0,0.0,0.0}
\pgfusepath{stroke}
\end{pgfscope}
\begin{pgfscope}
\pgfsetlinewidth{3.0bp}
\pgfsetrectcap 
\pgfsetmiterjoin \pgfsetmiterlimit{10.0}
\pgfpathmoveto{\pgfpoint{345.5bp}{28.5bp}}
\pgfpathcurveto{\pgfpoint{345.5bp}{13.588318bp}}{\pgfpoint{308.563492bp}{1.5bp}}{\pgfpoint{263.0bp}{1.5bp}}
\pgfpathcurveto{\pgfpoint{217.436508bp}{1.5bp}}{\pgfpoint{180.5bp}{13.588318bp}}{\pgfpoint{180.5bp}{28.5bp}}
\pgfpathcurveto{\pgfpoint{180.5bp}{43.411682bp}}{\pgfpoint{217.436508bp}{55.5bp}}{\pgfpoint{263.0bp}{55.5bp}}
\pgfpathcurveto{\pgfpoint{308.563492bp}{55.5bp}}{\pgfpoint{345.5bp}{43.411682bp}}{\pgfpoint{345.5bp}{28.5bp}}
\pgfclosepath
\color[rgb]{0.0,0.0,0.0}
\pgfusepath{stroke}
\end{pgfscope}
\begin{pgfscope}
\pgfsetlinewidth{3.0bp}
\pgfsetrectcap 
\pgfsetmiterjoin \pgfsetmiterlimit{10.0}
\pgfpathmoveto{\pgfpoint{77.25bp}{84.5bp}}
\pgfpathlineto{\pgfpoint{77.25bp}{125.75bp}}
\pgfpathlineto{\pgfpoint{19.5bp}{125.75bp}}
\pgfpathlineto{\pgfpoint{19.5bp}{84.5bp}}
\pgfpathlineto{\pgfpoint{77.25bp}{84.5bp}}
\pgfclosepath
\color[rgb]{0.0,0.0,0.0}
\pgfusepath{stroke}
\end{pgfscope}
\begin{pgfscope}
\pgfsetlinewidth{3.0bp}
\pgfsetrectcap 
\pgfsetmiterjoin \pgfsetmiterlimit{10.0}
\pgfpathmoveto{\pgfpoint{338.75bp}{89.25bp}}
\pgfpathlineto{\pgfpoint{338.75bp}{130.5bp}}
\pgfpathlineto{\pgfpoint{281.0bp}{130.5bp}}
\pgfpathlineto{\pgfpoint{281.0bp}{89.25bp}}
\pgfpathlineto{\pgfpoint{338.75bp}{89.25bp}}
\pgfclosepath
\color[rgb]{0.0,0.0,0.0}
\pgfusepath{stroke}
\end{pgfscope}
\begin{pgfscope}
\pgfsetlinewidth{3.0bp}
\pgfsetrectcap 
\pgfsetmiterjoin \pgfsetmiterlimit{10.0}
\pgfpathmoveto{\pgfpoint{297.0bp}{182.75bp}}
\pgfpathlineto{\pgfpoint{313.5bp}{131.75bp}}
\color[rgb]{0.0,0.0,0.0}
\pgfusepath{stroke}
\end{pgfscope}
\begin{pgfscope}
\pgfsetlinewidth{3.0bp}
\pgfsetrectcap 
\pgfsetmiterjoin \pgfsetmiterlimit{10.0}
\pgfpathmoveto{\pgfpoint{42.0bp}{126.5bp}}
\pgfpathlineto{\pgfpoint{50.25bp}{179.75bp}}
\color[rgb]{0.0,0.0,0.0}
\pgfusepath{stroke}
\end{pgfscope}
\begin{pgfscope}
\pgfsetlinewidth{3.0bp}
\pgfsetrectcap 
\pgfsetmiterjoin \pgfsetmiterlimit{10.0}
\pgfpathmoveto{\pgfpoint{81.0bp}{190.25bp}}
\pgfpathlineto{\pgfpoint{127.5bp}{136.25bp}}
\color[rgb]{0.0,0.0,0.0}
\pgfusepath{stroke}
\end{pgfscope}
\begin{pgfscope}
\pgfsetlinewidth{3.0bp}
\pgfsetrectcap 
\pgfsetmiterjoin \pgfsetmiterlimit{10.0}
\pgfpathmoveto{\pgfpoint{265.5bp}{184.25bp}}
\pgfpathlineto{\pgfpoint{240.75bp}{137.75bp}}
\color[rgb]{0.0,0.0,0.0}
\pgfusepath{stroke}
\end{pgfscope}
\begin{pgfscope}
\pgfsetlinewidth{3.0bp}
\pgfsetrectcap 
\pgfsetmiterjoin \pgfsetmiterlimit{10.0}
\pgfpathmoveto{\pgfpoint{164.25bp}{134.0bp}}
\pgfpathlineto{\pgfpoint{246.75bp}{194.75bp}}
\color[rgb]{0.0,0.0,0.0}
\pgfusepath{stroke}
\end{pgfscope}
\begin{pgfscope}
\pgfsetlinewidth{3.0bp}
\pgfsetrectcap 
\pgfsetmiterjoin \pgfsetmiterlimit{10.0}
\pgfpathmoveto{\pgfpoint{87.75bp}{59.75bp}}
\pgfpathlineto{\pgfpoint{126.0bp}{74.75bp}}
\color[rgb]{0.0,0.0,0.0}
\pgfusepath{stroke}
\end{pgfscope}
\begin{pgfscope}
\pgfsetlinewidth{3.0bp}
\pgfsetrectcap 
\pgfsetmiterjoin \pgfsetmiterlimit{10.0}
\pgfpathmoveto{\pgfpoint{233.25bp}{73.25bp}}
\pgfpathlineto{\pgfpoint{261.75bp}{56.0bp}}
\color[rgb]{0.0,0.0,0.0}
\pgfusepath{stroke}
\end{pgfscope}
\begin{pgfscope}
\pgftransformcm{1.0}{0.0}{0.0}{1.0}{\pgfpoint{251.25bp}{216.5bp}}
\pgftext[left,base]{Cancer  $Y_i$}
\end{pgfscope}
\begin{pgfscope}
\pgftransformcm{1.0}{0.0}{0.0}{1.0}{\pgfpoint{30.0bp}{211.25bp}}
\pgftext[left,base]{Smoking $Z_{ik}$}
\end{pgfscope}
\begin{pgfscope}
\pgftransformcm{1.0}{0.0}{0.0}{1.0}{\pgfpoint{124.5bp}{115.25bp}}
\pgftext[left,base]{Smoking}
\end{pgfscope}
\begin{pgfscope}
\pgftransformcm{1.0}{0.0}{0.0}{1.0}{\pgfpoint{124.25bp}{103.0bp}}
\pgftext[left,base]{Risk}
\end{pgfscope}
\begin{pgfscope}
\pgftransformcm{1.0}{0.0}{0.0}{1.0}{\pgfpoint{124.25bp}{91.0bp}}
\pgftext[left,base]{$\rho_i$}
\end{pgfscope}
\begin{pgfscope}
\pgftransformcm{1.0}{0.0}{0.0}{1.0}{\pgfpoint{216.75bp}{115.25bp}}
\pgftext[left,base]{Cancer}
\end{pgfscope}
\begin{pgfscope}
\pgftransformcm{1.0}{0.0}{0.0}{1.0}{\pgfpoint{217.5bp}{103.75bp}}
\pgftext[left,base]{Risk}
\end{pgfscope}
\begin{pgfscope}
\pgftransformcm{1.0}{0.0}{0.0}{1.0}{\pgfpoint{224.25bp}{93.25bp}}
\pgftext[left,base]{$U_i$}
\end{pgfscope}
\begin{pgfscope}
\pgftransformcm{1.0}{0.0}{0.0}{1.0}{\pgfpoint{32.25bp}{109.25bp}}
\pgftext[left,base]{Income}
\end{pgfscope}
\begin{pgfscope}
\pgftransformcm{1.0}{0.0}{0.0}{1.0}{\pgfpoint{32.25bp}{97.0bp}}
\pgftext[left,base]{$X_i$}
\end{pgfscope}
\begin{pgfscope}
\pgftransformcm{1.0}{0.0}{0.0}{1.0}{\pgfpoint{294.75bp}{114.5bp}}
\pgftext[left,base]{Pollution}
\end{pgfscope}
\begin{pgfscope}
\pgftransformcm{1.0}{0.0}{0.0}{1.0}{\pgfpoint{295.25bp}{102.0bp}}
\pgftext[left,base]{$P_i$}
\end{pgfscope}
\begin{pgfscope}
\pgftransformcm{1.0}{0.0}{0.0}{1.0}{\pgfpoint{28.5bp}{28.25bp}}
\pgftext[left,base]{Neighbouring smoking}
\end{pgfscope}
\begin{pgfscope}
\pgftransformcm{1.0}{0.0}{0.0}{1.0}{\pgfpoint{28.25bp}{16.0bp}}
\pgftext[left,base]{$W_j | j \sim i$ }
\end{pgfscope}
\begin{pgfscope}
\pgftransformcm{1.0}{0.0}{0.0}{1.0}{\pgfpoint{209.25bp}{36.5bp}}
\pgftext[left,base]{Neighbouring Cancer}
\end{pgfscope}
\begin{pgfscope}
\pgftransformcm{1.0}{0.0}{0.0}{1.0}{\pgfpoint{209.25bp}{24.0bp}}
\pgftext[left,base]{$U_j | j \sim i$}
\end{pgfscope}
\end{pgfpicture}