### 2009-08-21 docs/hamilton/slides.rnw

\documentclass[xcolor=dvipsnames]{beamer}

\usetheme{Singapore}
\usecolortheme[named=RawSienna]{structure}

\usefonttheme{serif}
\setbeamertemplate{navigation symbols}{}
\setbeamertemplate{footline}[frame number] 
\usepackage[nogin]{c:/brownpe/share/Sweave}
\title{Beyond disease mapping: models for disease incidence data with spatial risk factors}
\author{Patrick Brown\\ Billy Chang, Susitha Wanigaratne}
\institute{Cancer Care Ontario\\ Dalla Lana School of Public Health, University of Toronto}

\usepackage{amsmath,tikz}
\graphicspath{{Figures/}}
\setkeys{Gin}{width=\textwidth}
\SweaveOpts{echo=FALSE,fig=true,prefix.string=Figures/G,height=2,width=3}

\begin{document}

\frame{\maketitle}

\begin{frame}
\frametitle{Hamilton, Canada}
\includegraphics[trim=1cm 12cm 1cm 6cm,clip=true]{hamiltonMap.pdf}

\includegraphics{YorkBlvdHamiltonA.JPG}
\end{frame}

\begin{frame}

\frametitle{Lung cancer in Hamilton}
% left bottom right top

\includegraphics[trim=10cm 6cm 9cm 10cm,clip=true,width=0.75\textwidth]{Female-CancerExpected.pdf}
~
\includegraphics[trim=21cm 8cm 4cm 8cm,clip=true,width=0.18\textwidth]{Female-CancerExpected.pdf}
\end{frame}

\begin{frame}
\frametitle{Smoking in Hamilton}
\includegraphics[trim=10cm 6cm 9cm 10cm,clip=true,width=0.75\textwidth]{Female-eversmoking.pdf}
~
\includegraphics[trim=21cm 8cm 4cm 8cm,clip=true,width=0.18\textwidth]{Female-eversmoking.pdf}
\end{frame}

\begin{frame}
\frametitle{The Problem}
\begin{itemize}
	\item Create a model for predicting cancer rates in Hamilton, allowing for low counts in small areas
	\item Allow explanatory variables (smoking)
	\item But how do we get smoking rates?

\begin{itemize}
	\item And how well can we predict them?
\end{itemize}
\end{itemize}
\end{frame}

\begin{frame}
\frametitle{Issues with small area maps}

\begin{block}{Small counts}
	mostly 0 or 1
\begin{itemize}
\item The usual estimates of incidence rates perform poorly
	\item Methods based on the normal-distribution aren't appropriate
\end{itemize}

\end{block}
\begin{block}{Spatial dependence}
With regions close together being more similar than regions further apart
\begin{itemize}
	\item The usual regression models underestimate standard errors
  \item but this helps us deal with small counts
\end{itemize}

\end{block}

\end{frame}

\begin{frame}
\frametitle{This talk}
will accomplish the following:
\begin{itemize}
\item  present a model for mapping disease risk in small areas with low case counts

\item  show how to incorporate risk factors observed imprecisely, like smoking rates

\item  apply these methods to mapping cancer risk in Hamilton at the DA level, adjusting for smoking rates.\end{itemize}
\end{frame}

\begin{frame}
\frametitle{Cancer Data}
For each census area $i$ we have
\begin{itemize}
	\item Expected cancer case count $E_i$

\begin{itemize}
\item $E_i = C_i \theta$
	\item $C_i$ is the (vector) of populations by 5-year age group
	\item $\theta$ are the (assumed known) rates by age group
\end{itemize}
\item Count of cancer cases $Y_i$, often 0 or 1
\item Estimated pollution levels $P_i$
\end{itemize}
\end{frame}

\begin{frame}
\frametitle{Smoking data}
From the Canadian Community Household Survey (CCHS), we have
\begin{itemize}
	\item $N_i$ individuals surveyed in region $i$, often 0.
	\item The age $A_{ij}$ and smoking status $Z_{ij}$ of individual $j$ in region $i$
	\item The average household income $X_i$ in the region.
\end{itemize}

\end{frame}

\begin{frame}
\frametitle{Smoking model}
\vspace*{-0.5cm}
\begin{align*}
pr(Z_{ij} =1) = & \pi_{ij}\\
 \text{logit}(\pi_{ij}) = & f(A_{ij};\psi) + \rho_i\\
\rho_i = & X_i \alpha + W_i\\
 W_i \sim & \text{BYM}(\sigma_1, \sigma_2)
\end{align*}
\vspace*{-0.5cm}
\begin{itemize}
\item $f(a;\psi)$ is a smoothly-varying function of age
\begin{itemize}
	\item $\psi$ are the coefficients on linear splines with fixed knots
\end{itemize}
	\item The smoking risk surface $\rho_i$  is unknown.  
	\item It depends on income $X_i$ and residual spatial variation $W_i$
	\item $W_i$ is a Besag, York and Mollie model, which is smoothly varying in space

\begin{itemize}
	\item $W_i$ depends on the values $W_k$ in neighbouring regions $k$
	\item Technically, it's the sum of a Markov random field and a spatially independent process
	\item $\sigma_1$ is the variance of the spatially smooth part of $W_i$, and $\sigma_2$ is the variance of the rough part.
\end{itemize}
\end{itemize}

\end{frame}

\begin{frame}
\frametitle{Dealing with small cancer counts}

\begin{columns}
\column{0.5\textwidth}
\begin{block}{The Model}
\[
Y_i \sim \text{Poisson}(E_i\lambda_i)
\]
\begin{itemize}
	\item $Y_i$ is the number of cases in region $i$
	\item $E_i$ is the expected count
	\item $\lambda_i$ is the relative risk
\end{itemize}
\end{block}
\column{0.5\textwidth}
\begin{block}{The Poisson Distriution}
<<poisson,fig=true,echo=false>>=
options(SweaveHooks=list(fig=function() par(mar=c(2.5,2.5,0.1,0.1), mgp=c(1.5, 0.5, 0), cex=0.6)))
xseq = 0:14
toplot = dpois(xseq, 0.5)
names(toplot) = as.character(xseq)
toplot2 = dpois(xseq, 1.5)
barplot(toplot, width=0.4, xlim = range(xseq), space=4, xlab='count', ylab='pr(count)',
	col='red',border=NA)
barplot(toplot, width=0.21, add=T, space=8.8, col='white', xlim = range(xseq),border=NA, axes=F, axisnames=F)

barplot(toplot2, width=0.2, add=T, space=9, col='blue', xlim = range(xseq),border=NA, axes=F)
@

mean = \textcolor{red}{0.5},  \textcolor{blue}{1.5} 
\end{block}

\end{columns}

\begin{block}{Interpretation}
\begin{itemize}
	\item We don't observe the true rate $\lambda_i$
	\item but rather a noisy version of it in the form of case counts
	\item The Poisson distribution can cope with small counts and lots of zeros
\end{itemize}
\end{block}

\end{frame}

\begin{frame}
\frametitle{Cancer model}
\begin{align*}
Y_i \sim & \text{Poisson}(\lambda_i E_i)\\
\log(\lambda_{i}) = &   P_i \beta + \rho_i \gamma + U_i\\
 U_i \sim & \text{BYM}(\tau_1, \tau_2)
\end{align*}
\begin{itemize}
	\item The cancer relative risk surface $\lambda_i$  is unknown.  
	\item It depends on pollution $P_i$, smoking $\rho_i$ and residual spatial variation $U_i$
	\item $U_i$ is a spatially-smooth Besag, York and Mollie process which depends on it's neighbours
\item Similar models used by Haining et al (2007) and Best et al.
\end{itemize}

\end{frame}

\begin{frame}
\hspace*{-1cm}\input{graph.tex}
\end{frame}

\begin{frame}
\frametitle{Model fitting}
\begin{itemize}
	\item The data are not Normally-distributed, and there are  unobserved random effects.
	\item Uncertainty in smoking should carry over into the cancer model
	\item This makes the usual least squares or maximum likelihood methods problematic
	\item Bayesian inference using Markov Chain Monte Carlo simulation
	\item WinBUGS, with help from the R package R2WinBUGS and glmmBUGS
\end{itemize}
\end{frame}

\begin{frame}
\frametitle{Bayesian inference in a nutshell}
\begin{itemize}
	\item The models give the distributions

\begin{itemize}
	\item Cancer: $[Y | P, \rho, U, \beta, \gamma]$
	\item Smoking: $[Z  | X, W, \alpha]$
	\item Spatial surfaces $[U|\tau]$, $[W|\sigma]$
	\item Priors $[\gamma]$,$[\beta]$, $[\alpha]$ $[\sigma]$
$[\tau]$
\end{itemize}
\item Given the data $Y$ and $Z$, what are the distributions of the surfaces
\begin{itemize}
	\item Smoking: $[W|Y,Z]$
	\item Cancer: $[U|Y,Z]$
\end{itemize}
\item MCMC draws a number of equally likely values from these distributions
\item The average of these values is our best guess at the risk surfaces
\end{itemize}

\end{frame}

\begin{frame}
\frametitle{Adjusted risk surfaces}
\begin{enumerate}
	\item Cancer relative risk surface:
	\[\lambda_{i} =   \exp(P_i \beta) \exp( \rho_i \gamma)\exp( U_i)\]
	a value of $\lambda_4=1.3$ means an individual living in region 4 has a 30\% higher risk than the Ontario average for their age group 
	\item Remove the effect of smoking:
	\[\lambda_{i}/ \exp( \rho_i \gamma) = \exp(P_i \beta) \exp( U_i)\]
	A 1.3 means that compared to a typical region in Ontario with the same smoking rate region 4 but average pollution, risk is 30\% higher.

	\item Remove the effect of smoking and pollution:
	\[\lambda_{i}/ \exp( \rho_i \gamma)\exp(P_i \beta) =  \exp( U_i)\]
	Compared to a typical region the same smoking rate and pollution as region 4, risk is 30\% higher.
\end{enumerate}
\end{frame}

\begin{frame}

\frametitle{Cancer risk}
% left bottom right top

\includegraphics[trim=10cm 6cm 9cm 10cm,clip=true,width=0.75\textwidth]{Female-CancerExpected.pdf}
~
\includegraphics[trim=21cm 8cm 4cm 8cm,clip=true,width=0.18\textwidth]{Female-CancerExpected.pdf}
\end{frame}

\begin{frame} 
\frametitle{Cancer risk, removing smoking effect}
\includegraphics[trim=10cm 6cm 9cm 10cm,clip=true,width=0.75\textwidth]{FemalesCancerRandomEffectPolllution.pdf}
~
\includegraphics[trim=21cm 8cm 4cm 8cm,clip=true,width=0.18\textwidth]{FemalesCancerRandomEffectPolllution.pdf}
\end{frame}

\begin{frame}
\frametitle{Cancer risk, removing smoking and pollution}
\includegraphics[trim=10cm 6cm 9cm 10cm,clip=true,width=0.75\textwidth]{FemalesCancerRandomEffect.pdf}
~
\includegraphics[trim=21cm 8cm 4cm 8cm,clip=true,width=0.18\textwidth]{FemalesCancerRandomEffect.pdf}
\end{frame}

\begin{frame}
\frametitle{Parameter estimates}

\end{frame}

\begin{frame}
\frametitle{Chain convergence}

\end{frame}

\begin{frame}
\frametitle{Limitations and Extensions}
\begin{description}
	\item[Continuous space]:  $\lambda(s)$ instead of $\lambda_i$
\begin{itemize}
	\item The real risk surface is continuous
	\item We observed spatially aggregated counts.
\end{itemize}
	\item[Changes in time]: $\lambda(s,t)$

\begin{itemize}
	\item To gain more power, use data from a long time period
	\item Allow for possible changes in risk over time
	\item Changing census boundaries are a problem!
\end{itemize}
\item[]
\end{description}
\end{frame}

\begin{frame}
\frametitle{Extensions continued}
\begin{description}
\item[Better smoking model]: make better use of survey data
\begin{itemize}
	\item Individuals are either current, former, or never smokers
	\item Rates at which starting and quitting happens vary with age
	\item We have cross-sectional, not longitudinal, information
	\item Multi-state model with time varying rates
	\item Spatially-varying start rate
\end{itemize}

\item[Uncertainty in pollution]:
\begin{itemize}
	\item Monitoring data are available
	\item pollution is another indirectly observed covariate?
\end{itemize}
\end{description}
\end{frame}

\begin{frame}
\frametitle{Conclusion}

\begin{itemize}
	\item Small area disease mapping in space and time is possible even when counts are low 
	\item Include covariates in the model to estimate the effect of potential risk factors
	\item It's possible to include covariates which are indirectly or imprecisely observed.
	\item So always adjust for smoking when mapping diseases?
\end{itemize}
\end{frame}

\end{document}