### 2013-11-18 pkg/mapmisc/inst/doc/mapmisc.rnw
changed results of colourScale to plot, col, colOpacity, also made methods for factors, rasters
git-svn-id: svn://svn.r-forge.r-project.org/svnroot/diseasemapping@629 edb9625f-4e0d-4859-8d74-9fd3b1da38cb

\documentclass[12pt]{article}

%\VignetteEngine{knitr::knitr}
%\VignetteIndexEntry{Overview of mapping with mapmisc}

\usepackage[margin=1in,landscape]{geometry}
\usepackage{graphicx,hyperref}
\DeclareGraphicsExtensions{.png}

\usepackage{subcaption,float}
\newcommand{\subfloat}[2][need a sub-caption]{\subcaptionbox{#1}{#2}}

\title{The mapmisc package}
\author{Patrick Brown}

\begin{document}
\maketitle

<<setup,echo=FALSE,cache=FALSE,results='hide',message=FALSE>>=
library('knitr')
opts_chunk$set(echo=TRUE,message=FALSE,
		dev='png',
		fig.align='center',fig.pos='H',width=80,
		fig.height=5,fig.width=3,out.width="0.3\\textwidth"
)
par(bg="white")
library(mapmisc)
library(rgdal)
@

\section{Introduction}

This package provides a few utilities for making nice maps in R, with an
emphasis on enabling maps in short tidy code chunks which are suitable for
Sweave and knitr documents.  The package duplicates the capabilities of
packages such as classInt, geonames, and OpenStreetMap, and the price of having
tidier code is much of the flexibility from these other packages has been lost
here.

The meuse data from the sp package
<<meusedata,echo=TRUE>>=
library('mapmisc')
library(sp)
data(meuse)
coordinates(meuse) <- ~x+y
proj4string(meuse) <- CRS("+init=epsg:28992")
class(meuse)
@

Get elevation data with
<<elevation,eval=FALSE>>=
meuseLL = spTransform(meuse, CRS("+init=epsg:4326"))
getData("SRTM", lon=xmin(extent(meuseLL)), 
	lat=ymin(extent(meuseLL)),path=tempdir())
nldElev = raster(paste(tempdir(), "/", "srtm_38_02.tif", sep=""))
nldElev = crop(nldElev, extent(meuseLL))
nldElev = projectRaster(nldElev, crs=proj4string(meuse))
@
Or simply do
<<dataneth>>=
data(netherlands)
@

The elevation data is a Raster.
<<classElev>>=
class(nldElev)
nldElev = crop(nldElev, extend(extent(meuse), 1000))
@

\section{Downloading background maps and city locations}

Get a background map covering the extent of the meuse data
<<openmap,eval=FALSE>>=
nldTiles = openmap(meuse)
@
{\tt nldTiles} is a Raster with the sampe projection as {\tt meuse}
<<openmapInfo>>=
class(nldTiles)
proj4string(nldTiles)
proj4string(meuse)
@
Maps which can be downloaded are shown at
\url{http://diseasemapping.r-forge.r-project.org/openmap/}

{\tt GNcities} is a wrapper for the function of the same name in the geonames
package.

<<cities,eval=FALSE>>=
nldCities = GNcities(meuse, maxRows=6)
@
A SpatialPointsDataFrame, with same map projection.
<<citiesInfo>>=
class(nldCities)
names(nldCities)
proj4string(nldCities)
@

\section{Making maps}

The {\tt map.new} function sets up a map in the current plot window with the
correct limits and aspect ratio for the object supplied, and without margins or
white space. {\tt scaleBar} adds a scale and north arrow.  It uses the map
projection of the argument supplied to calculate distances and find north.

<<simpleMap,fig.cap="simple map", fig.subcap=c("data locations","cities","elevation")>>=

# plot the data locations
map.new(meuse)
plot(nldTiles, add=TRUE)
points(meuse,col="red", cex=0.3)
scaleBar(proj4string(meuse),pos="topleft", bg="white")

# plot city names
map.new(meuse)
plot(nldTiles, add=TRUE)
points(nldCities)
text(nldCities, labels=nldCities$name, pos=3)
scaleBar(proj4string(meuse),pos="topleft", bg="white")

# plot elevation
map.new(meuse, legendRight=TRUE)
plot(nldTiles, add=TRUE)
plot(nldElev,add=TRUE,col=terrain.colors(8),alpha=0.6,legend.mar=2, legend.line=0,)
scaleBar(meuse,pos="topleft",bg="white")
@

Reproject the data so it's roughly aligned with the x-axis.

<<reproj,cache=TRUE>>=
rotatedCRS = CRS("+proj=omerc +lat_0=51 +lonc=5.75 +alpha=-50.0000000 +gamma=0.0 +k=1.000000 +x_0=500000.000 +y_0=0.000 +ellps=WGS84	+units=m")
meuseRot = spTransform(meuse, rotatedCRS)
tilesRot = projectRaster(nldTiles, crs=rotatedCRS, method='ngb')
tilesRot@legend@colortable = nldTiles@legend@colortable
elevRot = projectRaster(nldElev, crs=rotatedCRS)
@

And create new plots

<<rotatedMap,fig.cap="Rotated map", fig.subcap=c("elevation","data"), fig.height=3,fig.width=4,out.width="0.48\\textwidth">>=
# first elevation
map.new(meuseRot)
plot(tilesRot, add=TRUE)
plot(elevRot,add=TRUE,alpha=0.5,col=terrain.colors(8), legend=FALSE)
points(nldCities)
text(nldCities, labels=nldCities$name, pos=3)
scaleBar(meuseRot,pos="topleft", bg="white")

# then data locations
map.new(meuseRot)
plot(tilesRot, add=TRUE)
points(meuseRot,col="red", cex=0.3)
scaleBar(proj4string(meuseRot), bg="white")
@

\section{Inset maps}

<<world,eval=FALSE>>=
world = openmap(extent(-10,30,40,60),path="osm-no-labels")
@

<<inset,fig.cap="Inset map", fig.subcap=c("elevation","rotated"), fig.height=3,fig.width=4,out.width="0.48\\textwidth">>=

# not rotated
map.new(meuse,legendRight=TRUE)
plot(nldTiles, add=TRUE)
points(meuse)
scaleBar(proj4string(meuse),pos="bottomright", bg="white")
insetMap(meuse, "topright",map=world)

# rotated
map.new(meuseRot)
plot(tilesRot, add=TRUE)
points(meuseRot,col="red", cex=0.3)
scaleBar(proj4string(meuseRot), bg="white")
insetMap(meuseRot, "bottomleft",map=world)
@

\section{Legends}

Create a colour scale for plotting copper  concentrations
<<meuseScale>>=
cuScale = colourScale(meuse$copper, breaks=5, style='equal', 
		opacity=0.8, dec=-1, firstBreak=0)
@
and elevation, with transparency decreasing as elevation increases.
<<elevScale>>=
elevScale = colourScale(nldElev, style='equal',
		breaks=6, col=terrain.colors,
		firstBreak=0, dec=-1,opacity=c(0.2, 0.9))
@
Soil type is a categorical variable, create a factor and create a colour scale
of unique values
<<soilScale>>=
meuse$soilFac = factor(meuse$soil, levels=c(1,2,3), 
		labels=c("Calcareous","Non-Calc's","Red Brick"))
soilScale = colourScale(meuse$soilFac, style="unique",
		col="Set2")
@

<<meuseLegends,fig.subcap=c("Copper","soil", "elevation"), fig.cap="Meuse data again">>=
map.new(meuse)
plot(nldTiles, add=TRUE)
plot(meuse, col=cuScale$plot,add=TRUE,pch=16)
scaleBar(meuse, pos="topleft",bg="white")
legendBreaks("bottomright",  breaks=cuScale,
		title="gals/firkin",bg="white",adj=c(0,-2))

map.new(meuse)
plot(nldTiles, add=TRUE)
plot(meuse, col=soilScale$plot,add=TRUE,pch=16)
scaleBar(meuse, pos="topleft",bg="white")
legendBreaks("bottomright", breaks=soilScale,
		title="soil type", cex=0.7,bg="white")

map.new(meuse)
plot(nldTiles, add=TRUE)
image(nldElev, breaks=elevScale$breaks, col=elevScale$colOpacity, 
		legend=FALSE,add=TRUE)
scaleBar(meuse, pos="topleft",bg="white")
legendBreaks("left", breaks=elevScale, title='Metres',bg="white")
@

\end{document}