### 2015-03-27 pkg/gmrf/vignettesStuff/sim.rnw
waiting for cholmod_solve2
git-svn-id: svn://svn.r-forge.r-project.org/svnroot/diseasemapping@1094 edb9625f-4e0d-4859-8d74-9fd3b1da38cb

\documentclass[12pt]{article}
\usepackage[margin=1in]{geometry}
\usepackage{caption,subcaption,float,amsmath,filecontents}
\newcommand{\subfloat}[2][need a sub-caption]{\subcaptionbox{#1}{#2}}

<<setup,include=FALSE>>=
library('knitr')

chapter=''
opts_chunk$set(echo=TRUE, 
		dev='pdf', marginsp=1,  
		fig.align='',fig.pos='H',
		fig.width=5,fig.height=4,out.width='\\textwidth',
		tidy=TRUE,tidy.opts=list(reindent.spaces=2,
				width.cutoff=45)
)

knit_hooks$set(
		marginsp = function(before, options, envir){	
			if(!before) return()
# use small margins				
			par(mar=c(1.5+0.9*options$marginsp,
							1.5+0.9*options$marginsp,0.2,0.2), 
					mgp=c(1.45, 0.45, 0),cex=1.25)
		}		
)

options(width=67)

@

\begin{document}

Hello $1^5=1^6$ therefore $5=6$.

Simulate some data
<<simData,cache=TRUE>>=
set.seed(0)
library(geostatsp)

myraster = squareRaster(raster(extent(0,8000,0,6000), ncols=40,nrows=30))

themodel = c(range=6*xres(myraster),shape=2,variance=900)

theU = RFsimulate(myraster,model=themodel)

thecov = myraster
values(thecov) = c(rep(1,ncell(thecov)/2),
		rep(4,ncell(thecov)/2))
beta.x=5
theY = theU + beta.x*thecov

theData = brick(theY,thecov)
names(theData) = c("y","x")

fracNugget = 1/2
nuggetSd = sqrt(themodel['variance']*fracNugget)
thedf = as.data.frame(theData)
thedf$intercept = 1
Xmat = as.matrix(thedf[,c("intercept","x")])
thedf$yNoise = rnorm(nrow(thedf),
		thedf$y,nuggetSd)

@

nearest neighbours
<<nnmat,cache=TRUE>>=
theNN = NNmat(theData)
@

fit the model
<<fit,cache=TRUE>>=

Sar = exp(seq(log(0.01),log(0.1),len=24))

source("../R/loglikGmrf.R")
source("../R/lgmrfm.R")
#source("../R/conditionalGmrf.R")

Sar2 =  seq((0.05),(0.4),len=24)
Snugget =  seq((0.1), (1),len=25)
maternShape = themodel['shape']

resVanilla = loglikGmrf(
		oneminusar=Sar2 ,
		propNugget = Snugget ,
		Yvec=thedf$yNoise,
		Xmat=Xmat,
		shape=maternShape,
		NN=theNN,mc.cores=4
)

@

make some plots
<<plots,out.width='0.45\\textwidth',fig.cap='the results',fig.subcap=c('L','second'),echo=FALSE,marginsp=TRUE>>=

dseq = rev(c(0,0.5, 1,2,4,8,20,100))
thecol = mapmisc::colourScale(resVanilla['logL.ml',,],
		breaks=max(resVanilla['logL.ml',,]) - dseq, 
		col='RdYlGn',style='fixed',rev=TRUE)

plot(range(resVanilla['propNugget',,1]), 
		range(resVanilla['oneminusar',1,]),type='n',
		xlab='tausq/xisq',ylab='1-phi')
.filled.contour(resVanilla['propNugget',,1], resVanilla['oneminusar',1,],resVanilla['logL.ml',,],
		col=thecol$col,levels=thecol$breaks
)

mapmisc::legendBreaks("topleft",col=thecol$col,breaks=dseq,
		inset=0.025)

plot(1:10)

@
\end{document}