### 2015-03-27 pkg/gmrf/vignettesStuff/so2Data.rnw

\documentclass[12pt]{article}
\usepackage[margin=1in]{geometry}
\usepackage{caption,subcaption,float,amsmath,filecontents}
\newcommand{\subfloat}[2][need a sub-caption]{\subcaptionbox{#1}{#2}}

\title{SO2 Data}

<<setup,include=FALSE>>=
library('knitr')

chapter=''
opts_chunk$set(echo=FALSE, 
		dev='pdf', marginsp=1,  
		fig.align='',fig.pos='H',
		fig.width=4,fig.height=3,out.width='\\textwidth',
		fig.path=file.path('Figures', chapter, ""),
		cache.path=file.path('cache', chapter, ""),
		tidy=TRUE,tidy.opts=list(reindent.spaces=2,code.width=45)
)

heavyCacheDate=NULL
lightCacheDate=NULL

knit_hooks$set(
		marginsp = function(before, options, envir){	
			if(!before) return()
# use small margins				
			par(mar=c(1.5+0.9*options$marginsp,
							1.5+0.9*options$marginsp,0.2,0.2), 
					mgp=c(1.45, 0.45, 0),cex=1.25)
		}		
)

library('RandomFields')
library('raster')
library('rgdal')
library('geostatsp')
library('mapmisc')
library('Hmisc')

evalP = Sys.info()['user'] =='patrick'

@

\begin{document}

<<child="/home/patrick/workspace/ssr/data/ozone.Rnw">>=
@

<<chinaDF>>=
china2 = crop(china, extent(80,100,20,35))
chinaDF = as.data.frame(china2, xy=TRUE)
chinaDF[is.na(chinaDF$pop),'pop'] = 0
chinaDF$logpop = log(chinaDF$pop+1)
chinaDF$popLow = pmin(0,chinaDF$logpop-1)
chinaDF$popHigh = pmax(0,chinaDF$logpop-1)
chinaDF$popVeryHigh = pmax(0,chinaDF$logpop-4)
chinaDF$latLow = pmin(0, chinaDF$y-45)
chinaDF$latHigh = pmax(0, chinaDF$y-45)
@

<<>>=
save(chinaDF,china2,file="chinaSo2.RData")
@

<<>>=
#plot(gam(y ~  s(lognight) + s(logpop,k=7,fx=T),data=bcAsiaDF))
@
\end{document}