### 2015-03-27 pkg/gmrf/vignettesStuff/v1.0.rnw

\documentclass[12pt]{article}
\usepackage[margin=1in]{geometry}
\usepackage{caption,float,amsmath,filecontents,subfig}

<<setup,include=FALSE>>=
library('knitr')

chapter=''
opts_chunk$set(echo=TRUE,
  	dev='pdf', marginsp=1,
		fig.align='',fig.pos='H',
		fig.width=5,fig.height=4,out.width='\\textwidth',
		tidy=TRUE,tidy.opts=list(reindent.spaces=2,
				width.cutoff=45)
)

knit_hooks$set(
		marginsp = function(before, options, envir){	
			if(!before) return()
# use small margins				
			par(mar=c(1.5+0.9*options$marginsp,
							1.5+0.9*options$marginsp,0.2,0.2),
					mgp=c(1.45, 0.45, 0),cex=1.25)
		}		
)

options(width=55)

@

\begin{document}
Make 10 raster datas, fit lgmrfm function
\begin{enumerate}

\item The base raster
<<Model1, cache=TRUE>>=
themodel = c(range=4*xres(myraster),shape=1,variance=900,cellSize = xres(myraster))

@
<<Sim_data1, cache=TRUE,echo=FALSE>>=
set.seed(1234)
library(geostatsp)
library(MASS)
myraster = squareRaster(raster(extent(0,8000,0,6000), ncols=40,nrows=30))

themodel = c(range=4*xres(myraster),shape=1,variance=900,cellSize = xres(myraster))

tr_oneminusar = 1-(as.numeric(themodel['range']^2/(themodel['range']^2+2*xres(myraster)^2*themodel['shape'])))
tr_xisq = themodel['variance']*4*pi*themodel['shape']*(4/(1-tr_oneminusar)-4)^(themodel['shape'])

theU_mat = RFsimulate(myraster,model=themodel)

myprec = maternGmrfPrec(myraster, param = themodel)
theU_gmrf = mvrnorm(mu = rep(0, dim(myprec)[1]), Sigma = tr_xisq*solve(myprec))

thecov = myraster
values(thecov) = c(rep(1,ncell(thecov)/2),
                   rep(4,ncell(thecov)/2))
beta.x=5
theY_mat = theU_mat + beta.x*thecov
theY_gmrf = theU_gmrf + beta.x*thecov

fracNugget = 1/3
nuggetSd = sqrt(themodel['variance']*fracNugget)
data_mat = brick(theY_mat,thecov)
data_gmrf = brick(theY_gmrf, thecov)
names(data_mat) = c("y","x")
names(data_gmrf) = c('y','x')
thedf = as.data.frame(data_mat)
thedf$yNoise = rnorm(nrow(thedf),thedf$y,nuggetSd)
theYN = myraster
values(theYN) = thedf$yNoise
theData_mat = brick(theYN,thecov)
names(theData_mat) = c("y","x")
theNN_mat = NNmat(theData_mat)

thedf = as.data.frame(data_gmrf)
thedf$yNoise = rnorm(nrow(thedf),thedf$y,nuggetSd)
theYN = myraster
values(theYN) = thedf$yNoise
theData_gmrf = brick(theYN,thecov)
names(theData_gmrf) = c("y","x")
theNN_gmrf = NNmat(theData_gmrf)

tr_propNugget = as.numeric(nuggetSd^2/tr_xisq)

Sar = exp(seq(log(0.05),log(0.3),len=24))
Snugget =  seq((0.01), (0.1),len=24)

source('E:/xbw/my docs/PERSONAL/Research/2014Summer-P.Brown/R-code/loglikGmrf.R')
source('E:/xbw/my docs/PERSONAL/Research/2014Summer-P.Brown/R-code/lgmrfm.R')
source('E:/xbw/my docs/PERSONAL/Research/2014Summer-P.Brown/R-code/conditionalGmrf.R')

resRangeNoEdge_mat  = lgmrfm(
       data=theData_mat,
       formula = y ~ x,
      oneminusar = Sar, nugget = Snugget,
       shape=themodel['shape'],
       NN = theNN_mat
)

resRangeEdge_mat = lgmrfm(
       data=theData_mat,
       formula = y ~ x,
      oneminusar = Sar, nugget = Snugget,
       shape=themodel['shape'],
       NN = theNN_mat,
       adjustEdges=TRUE
)

resRangeNoEdge_gmrf  = lgmrfm(
       data=theData_gmrf,
       formula = y ~ x,
      oneminusar = Sar, nugget = Snugget,
       shape=themodel['shape'],
       NN = theNN_gmrf
)

resRangeEdge_gmrf = lgmrfm(
       data=theData_gmrf,
       formula = y ~ x,
      oneminusar = Sar, nugget = Snugget,
       shape=themodel['shape'],
       NN = theNN_gmrf,
       adjustEdges=TRUE
)

@

Plot the result with and without Edge corrections 
<<plot1_mat,out.width='0.45\\textwidth',fig.cap='Model1 results using Matern simulation',fig.subcap=c('without Edge Correction','with Edge Correction'),echo=FALSE,marginsp=TRUE>>=
dseq = rev(c(0,0.5, 1,2,4,8,10,30))
thecol = mapmisc::colourScale(resRangeNoEdge_mat$complete$array['logL.ml',,],
                              breaks=max(resRangeNoEdge_mat$complete$array['logL.ml',,]) - dseq,
                              col='RdYlGn',style='fixed',rev=TRUE)

plot(range(resRangeNoEdge_mat$complete$propNugget),
     range(resRangeNoEdge_mat$complete$oneminusar),type='n',
     xlab='tausq/xisq',ylab='1-ar')
.filled.contour(sort(resRangeNoEdge_mat$complete$propNugget), 
                sort(resRangeNoEdge_mat$complete$oneminusar),
                resRangeNoEdge_mat$complete$array['logL.ml',,],
                col=thecol$col,levels=thecol$breaks
)
points(tr_propNugget, tr_oneminusar,pch = 16)
#mapmisc::legendBreaks("topleft",col=thecol$col,breaks=dseq)

thecol = mapmisc::colourScale(resRangeEdge_mat$complete$array['logL.ml',,],
                              breaks=max(resRangeEdge_mat$complete$array['logL.ml',,]) - dseq,
                              col='RdYlGn',style='fixed',rev=TRUE)

plot(range(resRangeEdge_mat$complete$propNugget),
     range(resRangeEdge_mat$complete$oneminusar),type='n',
     xlab='tausq/xisq',ylab='1-ar')
.filled.contour(sort(resRangeEdge_mat$complete$propNugget), sort(resRangeEdge_mat$complete$oneminusar)
,resRangeEdge_mat$complete$array['logL.ml',,],col=thecol$col,levels=thecol$breaks)
points(tr_propNugget, tr_oneminusar,pch = 16)
#mapmisc::legendBreaks("topleft",col=thecol$col,breaks=dseq)
@

<<plot1_gmrf,out.width='0.45\\textwidth',fig.cap='Model1 results using GMRF simulation',fig.subcap=c('without Edge Correction','with Edge Correction'),echo=FALSE,marginsp=TRUE>>=
dseq = rev(c(0,0.5, 1,2,4,8,10,30))
thecol = mapmisc::colourScale(resRangeNoEdge_gmrf$complete$array['logL.ml',,],
                              breaks=max(resRangeNoEdge_gmrf$complete$array['logL.ml',,]) - dseq,
                              col='RdYlGn',style='fixed',rev=TRUE)

plot(range(resRangeNoEdge_gmrf$complete$propNugget),
     range(resRangeNoEdge_gmrf$complete$oneminusar),type='n',
     xlab='tausq/xisq',ylab='1-ar')
.filled.contour(sort(resRangeNoEdge_gmrf$complete$propNugget), 
                sort(resRangeNoEdge_gmrf$complete$oneminusar),
                resRangeNoEdge_gmrf$complete$array['logL.ml',,],
                col=thecol$col,levels=thecol$breaks
)
points(tr_propNugget, tr_oneminusar,pch = 16)
#mapmisc::legendBreaks("topleft",col=thecol$col,breaks=dseq)

thecol = mapmisc::colourScale(resRangeEdge_gmrf$complete$array['logL.ml',,],
                              breaks=max(resRangeEdge_gmrf$complete$array['logL.ml',,]) - dseq,
                              col='RdYlGn',style='fixed',rev=TRUE)

plot(range(resRangeEdge_gmrf$complete$propNugget),
     range(resRangeEdge_gmrf$complete$oneminusar),type='n',
     xlab='tausq/xisq',ylab='1-ar')
.filled.contour(sort(resRangeEdge_gmrf$complete$propNugget), sort(resRangeEdge_gmrf$complete$oneminusar)
,resRangeEdge_gmrf$complete$array['logL.ml',,],col=thecol$col,levels=thecol$breaks)
points(tr_propNugget, tr_oneminusar,pch = 16)
#mapmisc::legendBreaks("topleft",col=thecol$col,breaks=dseq)

@

\item A bigger range
<<Model2, cache=TRUE>>=
themodel = c(range=8*xres(myraster),shape=1,variance=900,cellSize = xres(myraster))

@

 <<Sim_data2, cache=TRUE, echo=FALSE>>=
set.seed(1234)
library(geostatsp)
library(MASS)
myraster = squareRaster(raster(extent(0,8000,0,6000), ncols=40,nrows=30))

themodel = c(range=8*xres(myraster),shape=1,variance=900,cellSize = xres(myraster))

tr_oneminusar = 1-(as.numeric(themodel['range']^2/(themodel['range']^2+2*xres(myraster)^2*themodel['shape'])))
tr_xisq = themodel['variance']*4*pi*themodel['shape']*(4/(1-tr_oneminusar)-4)^(themodel['shape'])

theU_mat = RFsimulate(myraster,model=themodel)

myprec = maternGmrfPrec(myraster, param = themodel)
theU_gmrf = mvrnorm(mu = rep(0, dim(myprec)[1]), Sigma = tr_xisq*solve(myprec))

thecov = myraster
values(thecov) = c(rep(1,ncell(thecov)/2),
                   rep(4,ncell(thecov)/2))
beta.x=5
theY_mat = theU_mat + beta.x*thecov
theY_gmrf = theU_gmrf + beta.x*thecov

fracNugget = 1/3
nuggetSd = sqrt(themodel['variance']*fracNugget)
data_mat = brick(theY_mat,thecov)
data_gmrf = brick(theY_gmrf, thecov)
names(data_mat) = c("y","x")
names(data_gmrf) = c('y','x')
thedf = as.data.frame(data_mat)
thedf$yNoise = rnorm(nrow(thedf),thedf$y,nuggetSd)
theYN = myraster
values(theYN) = thedf$yNoise
theData_mat = brick(theYN,thecov)
names(theData_mat) = c("y","x")
theNN_mat = NNmat(theData_mat)

thedf = as.data.frame(data_gmrf)
thedf$yNoise = rnorm(nrow(thedf),thedf$y,nuggetSd)
theYN = myraster
values(theYN) = thedf$yNoise
theData_gmrf = brick(theYN,thecov)
names(theData_gmrf) = c("y","x")
theNN_gmrf = NNmat(theData_gmrf)

tr_propNugget = as.numeric(nuggetSd^2/tr_xisq)

Sar = exp(seq(log(0.01),log(0.07),len=24))
Snugget =  seq((0.1), (0.4),len=24)

source('E:/xbw/my docs/PERSONAL/Research/2014Summer-P.Brown/R-code/loglikGmrf.R')
source('E:/xbw/my docs/PERSONAL/Research/2014Summer-P.Brown/R-code/lgmrfm.R')
source('E:/xbw/my docs/PERSONAL/Research/2014Summer-P.Brown/R-code/conditionalGmrf.R')

resRangeNoEdge_mat  = lgmrfm(
       data=theData_mat,
       formula = y ~ x,
      oneminusar = Sar, nugget = Snugget,
       shape=themodel['shape'],
       NN = theNN_mat
)

resRangeEdge_mat = lgmrfm(
       data=theData_mat,
       formula = y ~ x,
      oneminusar = Sar, nugget = Snugget,
       shape=themodel['shape'],
       NN = theNN_mat,
       adjustEdges=TRUE
)

resRangeNoEdge_gmrf  = lgmrfm(
       data=theData_gmrf,
       formula = y ~ x,
      oneminusar = Sar, nugget = Snugget,
       shape=themodel['shape'],
       NN = theNN_gmrf
)

resRangeEdge_gmrf = lgmrfm(
       data=theData_gmrf,
       formula = y ~ x,
      oneminusar = Sar, nugget = Snugget,
       shape=themodel['shape'],
       NN = theNN_gmrf,
       adjustEdges=TRUE
)
 @

Plot the result with and without Edge corrections 
<<plot2_mat,out.width='0.45\\textwidth',fig.cap='Model2 results using Matern simulation',fig.subcap=c('without Edge Correction','with Edge Correction'),echo=FALSE,marginsp=TRUE>>=
dseq = rev(c(0,0.5, 1,2,4,8,10,30))
thecol = mapmisc::colourScale(resRangeNoEdge_mat$complete$array['logL.ml',,],
                              breaks=max(resRangeNoEdge_mat$complete$array['logL.ml',,]) - dseq,
                              col='RdYlGn',style='fixed',rev=TRUE)

plot(range(resRangeNoEdge_mat$complete$propNugget),
     range(resRangeNoEdge_mat$complete$oneminusar),type='n',
     xlab='tausq/xisq',ylab='1-ar')
.filled.contour(sort(resRangeNoEdge_mat$complete$propNugget), 
                sort(resRangeNoEdge_mat$complete$oneminusar),
                resRangeNoEdge_mat$complete$array['logL.ml',,],
                col=thecol$col,levels=thecol$breaks
)
points(tr_propNugget, tr_oneminusar,pch = 16)
#mapmisc::legendBreaks("topleft",col=thecol$col,breaks=dseq)

thecol = mapmisc::colourScale(resRangeEdge_mat$complete$array['logL.ml',,],
                              breaks=max(resRangeEdge_mat$complete$array['logL.ml',,]) - dseq,
                              col='RdYlGn',style='fixed',rev=TRUE)

plot(range(resRangeEdge_mat$complete$propNugget),
     range(resRangeEdge_mat$complete$oneminusar),type='n',
     xlab='tausq/xisq',ylab='1-ar')
.filled.contour(sort(resRangeEdge_mat$complete$propNugget), sort(resRangeEdge_mat$complete$oneminusar)
,resRangeEdge_mat$complete$array['logL.ml',,],col=thecol$col,levels=thecol$breaks)
points(tr_propNugget, tr_oneminusar,pch = 16)
#mapmisc::legendBreaks("topleft",col=thecol$col,breaks=dseq)
@

<<plot2_gmrf,out.width='0.45\\textwidth',fig.cap='Model2 results using GMRF simulation',fig.subcap=c('without Edge Correction','with Edge Correction'),echo=FALSE,marginsp=TRUE>>=
dseq = rev(c(0,0.5, 1,2,4,8,10,30))
thecol = mapmisc::colourScale(resRangeNoEdge_gmrf$complete$array['logL.ml',,],
                              breaks=max(resRangeNoEdge_gmrf$complete$array['logL.ml',,]) - dseq,
                              col='RdYlGn',style='fixed',rev=TRUE)

plot(range(resRangeNoEdge_gmrf$complete$propNugget),
     range(resRangeNoEdge_gmrf$complete$oneminusar),type='n',
     xlab='tausq/xisq',ylab='1-ar')
.filled.contour(sort(resRangeNoEdge_gmrf$complete$propNugget), 
                sort(resRangeNoEdge_gmrf$complete$oneminusar),
                resRangeNoEdge_gmrf$complete$array['logL.ml',,],
                col=thecol$col,levels=thecol$breaks
)
points(tr_propNugget, tr_oneminusar,pch = 16)
#mapmisc::legendBreaks("topleft",col=thecol$col,breaks=dseq)

thecol = mapmisc::colourScale(resRangeEdge_gmrf$complete$array['logL.ml',,],
                              breaks=max(resRangeEdge_gmrf$complete$array['logL.ml',,]) - dseq,
                              col='RdYlGn',style='fixed',rev=TRUE)

plot(range(resRangeEdge_gmrf$complete$propNugget),
     range(resRangeEdge_gmrf$complete$oneminusar),type='n',
     xlab='tausq/xisq',ylab='1-ar')
.filled.contour(sort(resRangeEdge_gmrf$complete$propNugget), sort(resRangeEdge_gmrf$complete$oneminusar)
,resRangeEdge_gmrf$complete$array['logL.ml',,],col=thecol$col,levels=thecol$breaks)
points(tr_propNugget, tr_oneminusar,pch = 16)
#mapmisc::legendBreaks("topleft",col=thecol$col,breaks=dseq)

@

\item A smaller variance
<<Model3, cache=TRUE>>=
themodel = c(range=4*xres(myraster),shape=1,variance=9,cellSize = xres(myraster))

@

 <<Sim_data3, cache=TRUE, echo=FALSE, warning=FALSE>>=
set.seed(1234)
library(geostatsp)
library(MASS)
myraster = squareRaster(raster(extent(0,8000,0,6000), ncols=40,nrows=30))

themodel = c(range=4*xres(myraster),shape=1,variance=9,cellSize = xres(myraster))

tr_oneminusar = 1-(as.numeric(themodel['range']^2/(themodel['range']^2+2*xres(myraster)^2*themodel['shape'])))
tr_xisq = themodel['variance']*4*pi*themodel['shape']*(4/(1-tr_oneminusar)-4)^(themodel['shape'])

theU_mat = RFsimulate(myraster,model=themodel)

myprec = maternGmrfPrec(myraster, param = themodel)
theU_gmrf = mvrnorm(mu = rep(0, dim(myprec)[1]), Sigma = tr_xisq*solve(myprec))

thecov = myraster
values(thecov) = c(rep(1,ncell(thecov)/2),
                   rep(4,ncell(thecov)/2))
beta.x=5
theY_mat = theU_mat + beta.x*thecov
theY_gmrf = theU_gmrf + beta.x*thecov

fracNugget = 1/3
nuggetSd = sqrt(themodel['variance']*fracNugget)
data_mat = brick(theY_mat,thecov)
data_gmrf = brick(theY_gmrf, thecov)
names(data_mat) = c("y","x")
names(data_gmrf) = c('y','x')
thedf = as.data.frame(data_mat)
thedf$yNoise = rnorm(nrow(thedf),thedf$y,nuggetSd)
theYN = myraster
values(theYN) = thedf$yNoise
theData_mat = brick(theYN,thecov)
names(theData_mat) = c("y","x")
theNN_mat = NNmat(theData_mat)

thedf = as.data.frame(data_gmrf)
thedf$yNoise = rnorm(nrow(thedf),thedf$y,nuggetSd)
theYN = myraster
values(theYN) = thedf$yNoise
theData_gmrf = brick(theYN,thecov)
names(theData_gmrf) = c("y","x")
theNN_gmrf = NNmat(theData_gmrf)

tr_propNugget = as.numeric(nuggetSd^2/tr_xisq)

Sar = exp(seq(log(0.05),log(0.25),len=24))
Snugget =  seq((0.025), (0.1),len=24)

source('E:/xbw/my docs/PERSONAL/Research/2014Summer-P.Brown/R-code/loglikGmrf.R')
source('E:/xbw/my docs/PERSONAL/Research/2014Summer-P.Brown/R-code/lgmrfm.R')
source('E:/xbw/my docs/PERSONAL/Research/2014Summer-P.Brown/R-code/conditionalGmrf.R')

resRangeNoEdge_mat  = lgmrfm(
       data=theData_mat,
       formula = y ~ x,
      oneminusar = Sar, nugget = Snugget,
       shape=themodel['shape'],
       NN = theNN_mat
)

resRangeEdge_mat = lgmrfm(
       data=theData_mat,
       formula = y ~ x,
      oneminusar = Sar, nugget = Snugget,
       shape=themodel['shape'],
       NN = theNN_mat,
       adjustEdges=TRUE
)

resRangeNoEdge_gmrf  = lgmrfm(
       data=theData_gmrf,
       formula = y ~ x,
      oneminusar = Sar, nugget = Snugget,
       shape=themodel['shape'],
       NN = theNN_gmrf
)

resRangeEdge_gmrf = lgmrfm(
       data=theData_gmrf,
       formula = y ~ x,
      oneminusar = Sar, nugget = Snugget,
       shape=themodel['shape'],
       NN = theNN_gmrf,
       adjustEdges=TRUE
)
 @

Plot the result with and without Edge corrections 
<<plot3_mat,out.width='0.45\\textwidth',fig.cap='Model3 results using Matern simulation',fig.subcap=c('without Edge Correction','with Edge Correction'),echo=FALSE,marginsp=TRUE>>=
dseq = rev(c(0,0.5, 1,2,4,8,10,30))
thecol = mapmisc::colourScale(resRangeNoEdge_mat$complete$array['logL.ml',,],
                              breaks=max(resRangeNoEdge_mat$complete$array['logL.ml',,]) - dseq,
                              col='RdYlGn',style='fixed',rev=TRUE)

plot(range(resRangeNoEdge_mat$complete$propNugget),
     range(resRangeNoEdge_mat$complete$oneminusar),type='n',
     xlab='tausq/xisq',ylab='1-ar')
.filled.contour(sort(resRangeNoEdge_mat$complete$propNugget), 
                sort(resRangeNoEdge_mat$complete$oneminusar),
                resRangeNoEdge_mat$complete$array['logL.ml',,],
                col=thecol$col,levels=thecol$breaks
)
points(tr_propNugget, tr_oneminusar,pch = 16)
#mapmisc::legendBreaks("topleft",col=thecol$col,breaks=dseq)

thecol = mapmisc::colourScale(resRangeEdge_mat$complete$array['logL.ml',,],
                              breaks=max(resRangeEdge_mat$complete$array['logL.ml',,]) - dseq,
                              col='RdYlGn',style='fixed',rev=TRUE)

plot(range(resRangeEdge_mat$complete$propNugget),
     range(resRangeEdge_mat$complete$oneminusar),type='n',
     xlab='tausq/xisq',ylab='1-ar')
.filled.contour(sort(resRangeEdge_mat$complete$propNugget), sort(resRangeEdge_mat$complete$oneminusar)
,resRangeEdge_mat$complete$array['logL.ml',,],col=thecol$col,levels=thecol$breaks)
points(tr_propNugget, tr_oneminusar,pch = 16)
#mapmisc::legendBreaks("topleft",col=thecol$col,breaks=dseq)
@

<<plot3_gmrf,out.width='0.45\\textwidth',fig.cap='Model3 results using GMRF simulation',fig.subcap=c('without Edge Correction','with Edge Correction'),echo=FALSE,marginsp=TRUE>>=
dseq = rev(c(0,0.5, 1,2,4,8,10,30))
thecol = mapmisc::colourScale(resRangeNoEdge_gmrf$complete$array['logL.ml',,],
                              breaks=max(resRangeNoEdge_gmrf$complete$array['logL.ml',,]) - dseq,
                              col='RdYlGn',style='fixed',rev=TRUE)

plot(range(resRangeNoEdge_gmrf$complete$propNugget),
     range(resRangeNoEdge_gmrf$complete$oneminusar),type='n',
     xlab='tausq/xisq',ylab='1-ar')
.filled.contour(sort(resRangeNoEdge_gmrf$complete$propNugget), 
                sort(resRangeNoEdge_gmrf$complete$oneminusar),
                resRangeNoEdge_gmrf$complete$array['logL.ml',,],
                col=thecol$col,levels=thecol$breaks
)
points(tr_propNugget, tr_oneminusar,pch = 16)
#mapmisc::legendBreaks("topleft",col=thecol$col,breaks=dseq)

thecol = mapmisc::colourScale(resRangeEdge_gmrf$complete$array['logL.ml',,],
                              breaks=max(resRangeEdge_gmrf$complete$array['logL.ml',,]) - dseq,
                              col='RdYlGn',style='fixed',rev=TRUE)

plot(range(resRangeEdge_gmrf$complete$propNugget),
     range(resRangeEdge_gmrf$complete$oneminusar),type='n',
     xlab='tausq/xisq',ylab='1-ar')
.filled.contour(sort(resRangeEdge_gmrf$complete$propNugget), sort(resRangeEdge_gmrf$complete$oneminusar)
,resRangeEdge_gmrf$complete$array['logL.ml',,],col=thecol$col,levels=thecol$breaks)
points(tr_propNugget, tr_oneminusar,pch = 16)
#mapmisc::legendBreaks("topleft",col=thecol$col,breaks=dseq)

@
\item A different cellSize
<<Sim_data4, cache=TRUE, echo=FALSE, warning=FALSE>>=
library(geostatsp)
library(MASS)
myraster = squareRaster(raster(extent(0,8000,0,6000), ncols=20,nrows=15))

themodel = c(range=4*xres(myraster),shape=1,variance=900,cellSize = xres(myraster))

tr_oneminusar = 1-(as.numeric(themodel['range']^2/(themodel['range']^2+2*xres(myraster)^2*themodel['shape'])))
tr_xisq = themodel['variance']*4*pi*themodel['shape']*(4/(1-tr_oneminusar)-4)^(themodel['shape'])

theU_mat = RFsimulate(myraster,model=themodel)

myprec = maternGmrfPrec(myraster, param = themodel)
theU_gmrf = mvrnorm(mu = rep(0, dim(myprec)[1]), Sigma = tr_xisq*solve(myprec))

thecov = myraster
values(thecov) = c(rep(1,ncell(thecov)/2),
                   rep(4,ncell(thecov)/2))
beta.x=5
theY_mat = theU_mat + beta.x*thecov
theY_gmrf = theU_gmrf + beta.x*thecov

fracNugget = 1/3
nuggetSd = sqrt(themodel['variance']*fracNugget)
data_mat = brick(theY_mat,thecov)
data_gmrf = brick(theY_gmrf, thecov)
names(data_mat) = c("y","x")
names(data_gmrf) = c('y','x')
thedf = as.data.frame(data_mat)
thedf$yNoise = rnorm(nrow(thedf),thedf$y,nuggetSd)
theYN = myraster
values(theYN) = thedf$yNoise
theData_mat = brick(theYN,thecov)
names(theData_mat) = c("y","x")
theNN_mat = NNmat(theData_mat)

thedf = as.data.frame(data_gmrf)
thedf$yNoise = rnorm(nrow(thedf),thedf$y,nuggetSd)
theYN = myraster
values(theYN) = thedf$yNoise
theData_gmrf = brick(theYN,thecov)
names(theData_gmrf) = c("y","x")
theNN_gmrf = NNmat(theData_gmrf)

tr_propNugget = as.numeric(nuggetSd^2/tr_xisq)

Sar = exp(seq(log(0.05),log(0.25),len=24))
Snugget =  seq((0.025), (0.1),len=24)

source('E:/xbw/my docs/PERSONAL/Research/2014Summer-P.Brown/R-code/loglikGmrf.R')
source('E:/xbw/my docs/PERSONAL/Research/2014Summer-P.Brown/R-code/lgmrfm.R')
source('E:/xbw/my docs/PERSONAL/Research/2014Summer-P.Brown/R-code/conditionalGmrf.R')

resRangeNoEdge_mat  = lgmrfm(
       data=theData_mat,
       formula = y ~ x,
      oneminusar = Sar, nugget = Snugget,
       shape=themodel['shape'],
       NN = theNN_mat
)

resRangeEdge_mat = lgmrfm(
       data=theData_mat,
       formula = y ~ x,
      oneminusar = Sar, nugget = Snugget,
       shape=themodel['shape'],
       NN = theNN_mat,
       adjustEdges=TRUE
)

resRangeNoEdge_gmrf  = lgmrfm(
       data=theData_gmrf,
       formula = y ~ x,
      oneminusar = Sar, nugget = Snugget,
       shape=themodel['shape'],
       NN = theNN_gmrf
)

resRangeEdge_gmrf = lgmrfm(
       data=theData_gmrf,
       formula = y ~ x,
      oneminusar = Sar, nugget = Snugget,
       shape=themodel['shape'],
       NN = theNN_gmrf,
       adjustEdges=TRUE
)
 @

Plot the result with and without Edge corrections 
<<plot4_mat,out.width='0.45\\textwidth',fig.cap='Model4 results using Matern simulation',fig.subcap=c('without Edge Correction','with Edge Correction'),echo=FALSE,marginsp=TRUE>>=
dseq = rev(c(0,0.5, 1,2,4,8,10,30))
thecol = mapmisc::colourScale(resRangeNoEdge_mat$complete$array['logL.ml',,],
                              breaks=max(resRangeNoEdge_mat$complete$array['logL.ml',,]) - dseq,
                              col='RdYlGn',style='fixed',rev=TRUE)

plot(range(resRangeNoEdge_mat$complete$propNugget),
     range(resRangeNoEdge_mat$complete$oneminusar),type='n',
     xlab='tausq/xisq',ylab='1-ar')
.filled.contour(sort(resRangeNoEdge_mat$complete$propNugget), 
                sort(resRangeNoEdge_mat$complete$oneminusar),
                resRangeNoEdge_mat$complete$array['logL.ml',,],
                col=thecol$col,levels=thecol$breaks
)
points(tr_propNugget, tr_oneminusar,pch = 16)
#mapmisc::legendBreaks("topleft",col=thecol$col,breaks=dseq)

thecol = mapmisc::colourScale(resRangeEdge_mat$complete$array['logL.ml',,],
                              breaks=max(resRangeEdge_mat$complete$array['logL.ml',,]) - dseq,
                              col='RdYlGn',style='fixed',rev=TRUE)

plot(range(resRangeEdge_mat$complete$propNugget),
     range(resRangeEdge_mat$complete$oneminusar),type='n',
     xlab='tausq/xisq',ylab='1-ar')
.filled.contour(sort(resRangeEdge_mat$complete$propNugget), sort(resRangeEdge_mat$complete$oneminusar)
,resRangeEdge_mat$complete$array['logL.ml',,],col=thecol$col,levels=thecol$breaks)
points(tr_propNugget, tr_oneminusar,pch = 16)
#mapmisc::legendBreaks("topleft",col=thecol$col,breaks=dseq)
@

<<plot4_gmrf,out.width='0.45\\textwidth',fig.cap='Model4 results using GMRF simulation',fig.subcap=c('without Edge Correction','with Edge Correction'),echo=FALSE,marginsp=TRUE>>=
dseq = rev(c(0,0.5, 1,2,4,8,10,30))
thecol = mapmisc::colourScale(resRangeNoEdge_gmrf$complete$array['logL.ml',,],
                              breaks=max(resRangeNoEdge_gmrf$complete$array['logL.ml',,]) - dseq,
                              col='RdYlGn',style='fixed',rev=TRUE)

plot(range(resRangeNoEdge_gmrf$complete$propNugget),
     range(resRangeNoEdge_gmrf$complete$oneminusar),type='n',
     xlab='tausq/xisq',ylab='1-ar')
.filled.contour(sort(resRangeNoEdge_gmrf$complete$propNugget), 
                sort(resRangeNoEdge_gmrf$complete$oneminusar),
                resRangeNoEdge_gmrf$complete$array['logL.ml',,],
                col=thecol$col,levels=thecol$breaks
)
points(tr_propNugget, tr_oneminusar,pch = 16)
#mapmisc::legendBreaks("topleft",col=thecol$col,breaks=dseq)

thecol = mapmisc::colourScale(resRangeEdge_gmrf$complete$array['logL.ml',,],
                              breaks=max(resRangeEdge_gmrf$complete$array['logL.ml',,]) - dseq,
                              col='RdYlGn',style='fixed',rev=TRUE)

plot(range(resRangeEdge_gmrf$complete$propNugget),
     range(resRangeEdge_gmrf$complete$oneminusar),type='n',
     xlab='tausq/xisq',ylab='1-ar')
.filled.contour(sort(resRangeEdge_gmrf$complete$propNugget), sort(resRangeEdge_gmrf$complete$oneminusar)
,resRangeEdge_gmrf$complete$array['logL.ml',,],col=thecol$col,levels=thecol$breaks)
points(tr_propNugget, tr_oneminusar,pch = 16)
#mapmisc::legendBreaks("topleft",col=thecol$col,breaks=dseq)

@
\end{enumerate}
\end{document}