### 2017-02-27 pkg/Pmisc/inst/docs/document.Rmd
git-svn-id: svn://svn.r-forge.r-project.org/svnroot/diseasemapping@1658 edb9625f-4e0d-4859-8d74-9fd3b1da38cb

```{r setup, include=FALSE}

if(any(commandArgs()=='mdToTex')) {
# if you'll be converting md to tex, can use the following
  knitr::knit_hooks$set(plot=knitr::hook_plot_tex)  
  knitr::opts_chunk$set(
   out.width='0.45\\textwidth', 
			dev='pdf', fig.align='center')
  mdToTex = TRUE
} else {
  knitr::knit_hooks$set(plot=Pmisc::hook_plot_mdsubfig)
  knitr::opts_chunk$set(dev='png', out.width='40%')
  mdToTex = FALSE
}

knitr::knit_hooks$set(
	margins = Pmisc::hook_plot_margins
)

knitr::opts_chunk$set(
  echo=FALSE, prompt=TRUE, comment=NA,
		margins=1, fig.cap=' ', fig.pos='ht',
		fig.width=2.75, fig.height=2,  dpi=250, 
		tidy=TRUE, tidy.opts=list(indent=2, width.cutoff=55)
)

options(width=80)

bibFile = 'myDocBib.bib'
cat('
@article{lindgren2011explicit,
  title={An explicit link between {G}aussian fields and {G}aussian {M}arkov random fields the stochastic partial differential equation approach},
  author={Lindgren, Finn and Rue, Haavard and Lindstrom, Johan},
  journal={Journal of the Royal Statistical Society B},
  volume={73},
  number={4},
  pages={423-498},
  year={2011},
  publisher={Wiley}
}
', file=bibFile, encoding='UTF8')

```

```{r theHeader, echo=FALSE}
Pmisc::markdownHeader(
	title="Example of a simple R markdown document with subfigures", 		
	author='Patrick Brown',
	bibliography = bibFile,
	geometry='margin=1in',
	csl = system.file('src','apa-no-doi-no-issue.csl', package='Pmisc'),
	css = system.file('src','article.css', package='Pmisc')
)
```

# Introduction

@lindgren2011explicit is a very good paper, but Figure [@fig:aplot] is dull.

$$
\begin{aligned}
f(x) = &x + \int_0^\infty \exp(-\theta u^2) du\\
\rho(x) = & \sum_i x^i
\end{aligned}
$$

```{r aplot, fig.ncol=2, fig.cap='Uninteresting plots', fig.subcap=c('stuff','qq','eeee','www')}
plot(1:10)

plot(1:10+2)

plot(1:10-4)

plot(1:10+4)
```

```{r anotherplot, fig.ncol=3, fig.cap='Other uninteresting plots', fig.subcap=month.name[1:6], out.width=c('30%', '0.3\\textwidth')[1+mdToTex]}

for(D in knitr::opts_current$get()$fig.subcap) {
  hist(runif(100), main='')
}
```

Below one will find Table [@tbl:theTable]

```{r theTable}

mytable = as.data.frame(matrix(runif(6),,3))
names(mytable) = month.name[1:ncol(mytable)]

Pmisc::mdTable(x=mytable, digits=3,
  caption = 'the table', caption.loc='bottom', 
  where = 'htb')

```

# References