### 2017-05-10 docs/survival/weibullcure.tex
git-svn-id: svn://svn.r-forge.r-project.org/svnroot/diseasemapping@1717 edb9625f-4e0d-4859-8d74-9fd3b1da38cb

\documentclass[a4paper,11pt]{article}
\usepackage[scale={0.8,0.9},centering,includeheadfoot]{geometry}
\usepackage{amstext}
\usepackage{amsmath}
\usepackage{verbatim}
\newcommand{\vect}[1]{\boldsymbol{#1}}
\begin{document}
\section*{Weibull With Cure Fraction}

\subsection*{Background}

\begin{align*}
Z_i & \sim \text{Bernoulli}(\rho)\\
[Y_i | Z_i = 0] &\sim \text{Weibull}(\lambda_i, \alpha)\\
[Y_i | Z_i = 1] & = \infty
\end{align*}

\subsection*{Parametrisation}

The Weibull is parametrized as \texttt{variant=0} of the \texttt{weibull} family.
\begin{displaymath}
    f(y) = 
    (1-\rho) \alpha y^{\alpha-1}
    \lambda\exp( - \lambda  y^{\alpha}), 
    \qquad  0 \leq y < \infty,
    \qquad \alpha>0, \qquad \lambda>0
\end{displaymath}

\begin{description}
\item[$\alpha$:] shape parameter.
\item[$\rho$:] the cure fraction parameter
\end{description}

\subsection*{Link-function}

The parameter $\lambda$ is linked to the linear predictor as:
\[
    \lambda = \exp(\eta)
\]

\subsection*{Hyperparameters}

The $\alpha$ parameter is represented as
\[
    \theta_1 = \log\alpha
\]
and $\rho$ is transformed to
\[
\theta_2 = \log[\rho / (1-\rho)].
\]
The priors are defined on $\theta$.

\subsection*{Specification}

Response variable $y$ must be given using $\texttt{inla.surv()}$

\subsubsection*{Hyperparameter spesification and default values}

\begin{quote}
    \input{../hyper/likelihood/weibullcure.tex}
%    \input{weibullcureHyper.tex}
\end{quote}

\subsection*{Example}

In the following example we estimate the parameters in a simulated
case \verbatiminput{example-weibullcure.R}

\subsection*{Notes}

\begin{itemize}
\item Weibull model can be used for right censored, left censored,
    interval censored data. If the observed times $y$ are large/huge,
    then this can cause numerical overflow in the likelihood routine.
    If you encounter this problem, try to scale the observatios,
    \verb|time = time / max(time)| or similar.
\end{itemize}

\end{document}

% LocalWords:  np Hyperparameters Ntrials

%%% Local Variables: 
%%% TeX-master: t
%%% End: