### 2018-02-05 pkg/mapmisc/inst/extdoc/world.Rmd
hopefully all the caching stuff works now
git-svn-id: svn://svn.r-forge.r-project.org/svnroot/diseasemapping@1903 edb9625f-4e0d-4859-8d74-9fd3b1da38cb

```{r theHeader, echo=FALSE}
knitr::knit_hooks$set(plot=knitr::hook_plot_tex) 
Pmisc::markdownHeader(
  title="Maps of the world", 		
  author='Patrick Brown',
  geometry='margin=1in'
)
```

```{r worldCrs, tidy=TRUE, warning=FALSE}
require('mapmisc', quietly=TRUE)

worldLL = openmap(extent(-180,180,-80,80), zoom=1, crs=mapmisc::crsLL)

Scentre = c('Grenwich'=0, Toronto=-79, 'Hong kong' = 114)

mollList = mapply(
    mapmisc::moll,     x= Scentre     )

worldMoll = mapply(     mapmisc::openmap,     x=mollList,     MoreArgs = list(zoom=1,
fact=2)     )

```

```{r worldMaps, fig.width=8, fig.height=3, fig.cap='World maps in different projections', fig.subcap = c('Long-lat', paste('Mollweide', names(mollList))), dev='png', out.width='0.8\\textwidth', warning=FALSE, fig.align='default'}
map.new(worldLL)
plot(worldLL, add=TRUE)

for(D in names(worldMoll)) {
  map.new(mollList[[D]])
  plot(worldMoll[[D]], add=TRUE)
}

```