### 2019-02-12 pkg/Pmisc/inst/docs/document.Rmd
subfigures fixed
git-svn-id: svn://svn.r-forge.r-project.org/svnroot/diseasemapping@2046 edb9625f-4e0d-4859-8d74-9fd3b1da38cb

```{r setup, include=FALSE}

if(any(commandArgs()=='mdToTex')) {
# if you'll be converting md to tex, can use the following
  knitr::knit_hooks$set(plot=knitr::hook_plot_tex)  
  knitr::opts_chunk$set(
    out.width='0.45\\textwidth', 
    dev='pdf', fig.align='center')
  mdToTex = TRUE
} else {
  knitr::knit_hooks$set(plot=Pmisc::hook_plot_mdsubfig)
  knitr::opts_chunk$set(dev='png')
  mdToTex = FALSE
}

knitr::knit_hooks$set(
  margins = Pmisc::hook_plot_margins
)

knitr::opts_chunk$set(
  echo=FALSE, prompt=TRUE, comment=NA,
  margins=1, fig.cap=' ', fig.pos='ht',
  fig.width=5.5, fig.height=4, 
  tidy=TRUE, tidy.opts=list(indent=2, width.cutoff=55)
)

options(width=80)

bibFile = 'myDocBib.bib'
cat('
    @article{lindgren2011explicit,
    title={An explicit link between {G}aussian fields and {G}aussian {M}arkov random fields the stochastic partial differential equation approach},
    author={Lindgren, Finn and Rue, Haavard and Lindstrom, Johan},
    journal={Journal of the Royal Statistical Society B},
    volume={73},
    number={4},
    pages={423-498},
    year={2011},
    publisher={Wiley}
    }
    ', file=bibFile, encoding='UTF8')

```

```{r theHeader, echo=FALSE}
Pmisc::markdownHeader(
  title="Example of a simple R markdown document with subfigures", 		
  author='Patrick Brown',
  bibliography = bibFile,
  geometry='margin=1in'
)

```

# Introduction {#sec:intro}

@lindgren2011explicit is a very good paper, but Figure [@fig:aplot] is dull.
There are some formulas in this document, from ([@eq:myeqn]) we can see
nothing useful.  If you are unsure of which Section we are in, be aware that
this is Section [@sec:intro].

$$
\begin{aligned}
f(x) = &x + \int_0^\infty \exp(-\theta u^2) du\\
\rho(x_1 \ldots x_N) = & \sum_i x^i
\end{aligned}
$$ {#eq:myeqn}

Now some plots.

```{r aplot, fig.ncol=2, fig.cap='Uninteresting plots', fig.subcap=c('stuff','qq','eeee','www')}
plot(1:10)

plot(1:10+2)

plot(1:10-4)

plot(1:10+4)
```

```{r anotherplot, fig.ncol=3, fig.cap='Other uninteresting plots', fig.subcap=month.name[1:6], out.width=Pmisc::out.width(0.32)}

for(D in knitr::opts_current$get()$fig.subcap) {
  hist(runif(100), main='')
}
```

Below one will find Table [@tbl:theTable]

```{r theTable}

mytable = as.data.frame(matrix(runif(6),,3))
names(mytable) = month.name[1:ncol(mytable)]

Pmisc::mdTable(x=mytable, digits=3,
  caption = 'the table', caption.loc='bottom', 
  where = 'htb')

```

# References