### 2023-12-06 pkg/mapmiscTerra/inst/extdoc/mapLayers.Rmd
maplayers list
git-svn-id: svn://svn.r-forge.r-project.org/svnroot/diseasemapping@3448 edb9625f-4e0d-4859-8d74-9fd3b1da38cb

<!--
%\VignetteEngine{knitr::knitr}
%\VignetteIndexEntry{Examples of map layers}
-->

# Examples of map layers

This document will be incomplete if there is no
internet connection when this document is compiled.  The full document is at
[diseasemapping.r-forge.r-project.org](http://diseasemapping.r-forge.r-project.org).

See [mc.bbbike.org/mc](http://mc.bbbike.org/mc/) for examples more map tiles 
(not all of which are compatible with `mapmisc`).

```{r setup, include=FALSE}
library('knitr')
library('mapmisc')

knitr::knit_hooks$set(plot=Pmisc::hook_plot_htmlsubfig)
knitr::opts_chunk$set(dev='png')

fact = 0.5
allTiles = FALSE

# uncomment for nicer images
# fact = 1 
# reRun = TRUE
# or run in sandbox	
# R --slave -e "knitr::knit('../pkg/mapmisc/vignettes/mapLayers.Rmd')" fact=1 allTiles=TRUE
# pandoc --to=html5 --output=mapLayers.html mapLayers.md
# cp -r figure ../www
# cp -r mapLayers.html ../www
theArgs = grep("^(fact|allTiles)=", commandArgs(), value=TRUE)
for(D in theArgs)
  eval(parse(text=D))

opts_chunk$set(dpi=50*fact, 
  dev.args=list(bg='white'))

cachePath = "/tmp/cacheLayers/"
dir.create(cachePath, showWarnings=FALSE)

theNames = names(osmTiles())

```

```{r theNames}
theNames
```

```{r theExtent, include=FALSE}
africaExtent = c(25, 25)
trafalgarSquare=c(-0.128056,51.508056)
stjohns=c(-52.683320, 47.566285)
```

```{r downloadTiles,include=FALSE}

  theNames = grep("sputnik", theNames, invert=TRUE, value=TRUE)
  print(theNames)
  print(cachePath)

  africaTiles = mapply(openmap, path=theNames,
    MoreArgs = list(x=africaExtent, zoom=2,
    verbose=TRUE, 
    crs = crsMerc, 
    cachePath = cachePath))

  londonTiles = mapply(openmap, path=theNames,
    MoreArgs = list(x=trafalgarSquare,zoom=12, 
    verbose=TRUE, 
    crs = crsMerc, 
    cachePath = cachePath))

  nfTiles = mapply(openmap, path=theNames,
    MoreArgs = list(x=stjohns, zoom=14,
    verbose=TRUE, 
    crs = crsMerc, 
    cachePath = cachePath))

```

The following are some examples of maps produced with `mymap=openmap(x, path="something")`

```{r plotSetup, include=FALSE}
mytiles = list(africa=africaTiles, london=londonTiles, newfoundland = nfTiles)
theCap = rep(theNames, each=length(mytiles))
```

```{r someTiles, echo=FALSE, fig.ncol=3, out.width='32%', fig.cap = 'some tiles', fig.height=4, fig.width=4, dpi=120, fig.subcap = theCap}

for(D in theNames){

# cat("\n\n# ", D, "\n\n")

      for(Dplot in names(mytiles)) {
        map.new(mytiles[[Dplot]][[D]])
        plot(mytiles[[Dplot]][[D]],add=TRUE, maxcells = 1e12)
    }
}
```

# Attributions

```{r attributionsExample, echo=TRUE}
openmapAttribution('osm', type='text', short=TRUE)
openmapAttribution('osm', type='text', short=FALSE)
```

```{r theAttributions, echo=FALSE, results='asis'}
for(D in theNames) {
  cat("## ", D, '\n\n')
    cat(openmapAttribution(D, 'html', short=TRUE), '\n\n')
    cat(openmapAttribution(D, 'html', short=FALSE), '\n\n')
}

```

```{r someTilesOld, eval=FALSE, echo=FALSE, results='asis', fig.height=5, fig.width=5}

for(D in theNames){
  Dsub = gsub("-",".", D)

  threeTiles = length(
    grep(
      paste("^", Dsub, "([rR]ed|[gG]reen|[bB]lue)$", sep=""), 
      names(mytiles[[1]])
    )
  ) == 3

  cat("\n\n# ", D, "\n\n")
  cat("`plot", "RGB"[threeTiles], 
    "(openmap(x,path=\"", D, "\"))`\n\n", 
    sep='')
  cat(openmapAttribution(D, 'html'), '\n\n')

  cat("\n|     |     |     |\n|-----|-----|-----|\n")
  for(Dplot in 1:length(mytiles)) {
    cat("|")

    oneTileName = grep(paste("^", Dsub, "$", sep=""), 
      names(mytiles[[Dplot]]), value=TRUE)
    threeTilesName = grep(paste("^", Dsub, 
        "([rR]ed|[gG]reen|[bB]lue)$", sep=""), 
      names(mytiles[[Dplot]]),value=TRUE)

    if(length(threeTilesName)==3) { # plotRGB
      map.new(mytiles[[Dplot]])

      plotRGB(mytiles[[Dplot]][[
          paste(Dsub, c("Red","Green","Blue"),sep="")
        ]],add=TRUE)

    } else if (length(oneTileName)==1){

      map.new(mytiles[[Dplot]])
      plot(mytiles[[Dplot]][[oneTileName]],add=TRUE)

    } else {

      plot(0.5, 0.8)
      text(0.5, 0.8, paste(D, "is\n not currently\navailable"))

    }

  }
  cat("|")

}

```